### 2021-01-17 paper/paper.tex
started writing paper in IEEE format

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Designing and Developing a Personalised Recommender System}

\author{\IEEEauthorblockN{Thomas Butterfield}
\IEEEauthorblockA{\textit{dept. Computer Science} \\
\textit{Durham University}\\
Durham, United Kingdom \\
thomas.butterfield@durham.ac.uk}
}

\maketitle

\section{Introduction}

\subsection{Domain of application}
This application is a hybrid recommender system using the Yelp dataset.

\subsection{Related work review}
Some related work

\subsection{Purpose/Aim}
The purpose and aim

\section{Methods}

\subsection{Data description}
Description of data

\subsection{Data preparation and future selection}
Prepare data

\subsection{Hybrid scheme}
Which two algorithms

\subsection{Recommendation techniques/algorithms}
What techniques

\subsection{Evaluation methods}
How to evaluate

\section{Implementation}

\subsection{Input interface}
What is the input interface

\subsection{Recommendation algorithm}
What algorithm

\subsection{Output interface}
What is the output interface

\section{Evaluation results}

\subsection{Comparison against baseline implementation}
Compare vs generic suggestions

\subsection{Comparison against hybrid recommenders in related studies}
Read some papers

\subsection{Ethical issues}
People's personal data

\section{Conclusion}

\subsection{Limitations}
What can it not do?

\subsection{Further developments}
What could I do in the future

\begin{thebibliography}{00}
\bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
\bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}

\end{document}

### 2021-01-17 code.py

@@ -78,9 +78,14 @@ q. Quit
                 exit()
     return choice
 
+def welcome():
+    print('Welcome to the hybrid recommender system using the Yelp dataset!')
+    print()
+
 
 if __name__ == '__main__':
-    print('Welcome to the hybrid recommender system using the Yelp dataset!')
+    welcome()
+    
     while True:
         choice = menu()
         if choice == 1:

### 2021-01-17 paper/paper.tex
filling out paper with initial thoughts

@@
 \subsection{Domain of application}
-This application is a hybrid recommender system using the Yelp dataset.
+The domain of this application is a recommender system for bars, using the Yelp dataset.
+Decide whether I want to define the time-span for the date, 
+depends on if there is too much data from all available time period.
+I am also taking into accounts Covid-19 data when generating recommendations.
 
@@
 \subsection{Purpose/Aim}
-The purpose and aim
+The purpose of this application is to give suitable suggestions for a user to go to.
+Using information about the user such as their location and personal preferences.
 
@@
 \subsection{Data description}
-Description of data
+The data I am using is taken from the Yelp dataset,
+it includes user reviews of different businesses and services in a specific location.
+These user reviews include ratings, text feedback, and other such information.
 
-\subsection{Data preparation and future selection}
+\subsection{Data preparation and feature selection}
 Prepare data
+The entire Yelp dataset is huge and much of it is not neccessary for my domain of bars.
+As such I prepared the data by eliminating any data not relevant to my domain.
+I selected features such as user ratings and a particular user's average rating,
+since if a user typically rates places they go highly, but rates a particular bar low,
+this is more significant than a user who always rates places low.
 
@@
 Which two algorithms
+A hybrid scheme is a good way to design a recommender system, since you can get the
+best of both algorithms if done properly.
+Meaning better recommendations than either algorithm could achieve individually.
 
 \subsection{Recommendation techniques/algorithms}
-What techniques
+The first recommender system is collaborative filtering.
+I am using a weighted mixed combination of these two systems in order to produce the results
+of my hybrid recommender system.
 
@@
 \subsection{Input interface}
-What is the input interface
+The input interface is the command line.
+The program offers users opportunites to input information about themselves,
+as well as make choices from a selection of items which the system provides or suggests.
 
@@
 \subsection{Output interface}
-What is the output interface
+The output interface is the command line.
+The system can output recommendations for bars which the user might like,
+as well as information about how the system works and why certain suggestions were made,
+at the user's request.
 

### 2021-01-29 paper/paper.tex
filtered business file to restaurants only

@@
 it includes user reviews of different businesses and services in a specific location.
+There are 10 cities included in the dataset: Montreal, Calgary, Toronto, Pittsburgh, Charlotte, 
+Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland. 
 These user reviews include ratings, text feedback, and other such information.
@@
 Prepare data
-The entire Yelp dataset is huge and much of it is not neccessary for my domain of bars.
+Of the 209,393 businesses in the dataset, 168,903 are still open.
+The entire Yelp dataset is huge and much of it is not necessary for my domain of bars.
 As such I prepared the data by eliminating any data not relevant to my domain.
@@
 \subsection{Further developments}
-What could I do in the future
+What could I do in the future \cite{5284958}.
 
-\begin{thebibliography}{00}
-\bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-\bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-\bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-\bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-\bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-\bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-\bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
-\vspace{12pt}
-\color{red}
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,references}
 

### 2021-01-29 code.py

@@ -82,9 +82,32 @@ def welcome():
     print('Welcome to the hybrid recommender system using the Yelp dataset!')
     print()
 
+def whichCity():
+    cities = ['Montreal','Calgary','Toronto','Pittsburgh','Charlotte','Urbana-Champaign','Phoenix','Las Vegas','Madison','Cleveland']
+    question = '''
+Which city are you closest to?
+    '''
+    for i in range(len(cities)):
+        question += '''
+{}. {}'''.format(i+1,cities[i])
+    print(question)
+    print()
+    check = False
+    while not check:
+        choice = input('Your choice > ')
+        try:
+            choice = int(choice)
+            if 1 <= choice <= len(cities):
+                check = True
+        except:
+            pass
+    return cities[choice-1]
+
 
 if __name__ == '__main__':
     welcome()
+    city = whichCity()
+    print(city)
     
     while True:
         choice = menu()
@@ -94,3 +117,15 @@ if __name__ == '__main__':
             pass
         elif choice == 3:
             info()
+
+# Resources:
+'''
+https://www.yelp.com/dataset
+https://www.yelp.com/dataset/documentation/faq
+https://www.yelp-support.com/Recommended_Reviews
+https://github.com/Yelp/dataset-examples/blob/master/json_to_csv_converter.py
+https://towardsdatascience.com/converting-yelp-dataset-to-csv-using-pandas-2a4c8f03bd88
+https://www.kaggle.com/yelp-dataset/yelp-dataset?select=yelp_academic_dataset_business.json
+https://github.com/Yelp/dataset-examples/issues/43
+https://surpriselib.com/
+'''

### 2021-01-29 filter.py

@@ -1,0 +1,83 @@
+
+import json
+import pandas as pd
+
+
+business_json_path = 'yelp_dataset/yelp_academic_dataset_business.json'
+'''
+totalCount = 0
+openCount = 0
+with open('yelp_dataset/yelp_academic_dataset_business.json','r') as f:
+    with open('processedData/business_filtered.json','w') as fout:
+        for line in f:
+            business = json.loads(line)
+            isOpen = business['is_open']
+            if isOpen == 1:
+                openCount += 1
+                fout.write(line)
+            totalCount += 1
+print(openCount)
+print(totalCount)
+'''
+
+df_b = pd.read_json(business_json_path, lines=True)
+
+df_b = df_b[df_b['is_open'] == 1]
+
+drop_columns = ['is_open']
+df_b = df_b.drop(drop_columns, axis=1)
+
+business_restaurant = df_b[df_b['categories'].str.contains('Restaurants',case=False,na=False)]
+
+df_explode = df_b.assign(categories = df_b.categories.str.split(', ')).explode('categories')
+
+print(df_explode.categories.value_counts())
+
+print(df_explode[df_explode['categories'].str.contains('Restaurants',case=True,na=False)].categories.value_counts())
+
+csv_name = 'processedData/business_restaurants.csv'
+business_restaurant.to_csv(csv_name, index=False)
+
+
+# filter by time
+# get number of reviews in last year?
+
+
+
+'''
+# filter dataset by city
+
+states = {}
+with open('yelp_dataset/yelp_academic_dataset_business.json') as f:
+    for line in f:
+        business = json.loads(line)
+        state = business['state']
+        city = business['city']
+        if state not in states:
+            states[state] = {city:1}
+        else:
+            if city not in states[state]:
+                states[state][city] = 1
+            else:
+                states[state][city] = states[state][city] + 1
+print(states)
+
+for state in states:
+    with open('processedData/'+state+'.txt','w') as fout:
+        for city in states[state]:
+            if states[state][city] > 1:
+                fout.write(city + '\n')
+
+
+
+print(data)
+states = []
+
+for business in data:
+    state = business['state']
+    if state not in states:
+        states.append(states)
+
+print(states)
+
+'''

### 2021-01-30 filter.py
filtered business json to csv restaurants only

@@ -29,15 +29,16 @@ df_b = df_b.drop(drop_columns, axis=1)
 
 business_restaurant = df_b[df_b['categories'].str.contains('Restaurants',case=False,na=False)]
 
-df_explode = df_b.assign(categories = df_b.categories.str.split(', ')).explode('categories')
+#df_explode = df_b.assign(categories = df_b.categories.str.split(', ')).explode('categories')
 
-print(df_explode.categories.value_counts())
+#print(df_explode.categories.value_counts())
 
-print(df_explode[df_explode['categories'].str.contains('Restaurants',case=True,na=False)].categories.value_counts())
+#print(df_explode[df_explode['categories'].str.contains('Restaurants',case=True,na=False)].categories.value_counts())
 
 csv_name = 'processedData/business_restaurants.csv'
 business_restaurant.to_csv(csv_name, index=False)
 
+# make dictionary of business IDs for restaurants only and filter tips,reviews, checkins for only those businesses
 
 # filter by time
 # get number of reviews in last year?

### 2021-02-03 paper/paper.tex
filtered business json to csv restaurants only
filtered dataset to restaurants only

@@
 Of the 209,393 businesses in the dataset, 168,903 are still open. 
+There are 43,980 restaurants which are open and contain 'Restaurant' as one of the categories. 
+502 businesses in the overall dataset don't list anything in their categories section. 
+43,052 businesses had at least 1 check-in. 
+There are 4,223,201 reviews of restaurants. 
+695,951 tips for these restaurants. 
+1,968,703 users. 
+43,992 restaurants listed with covid features, 8 restaurants which make up 12 duplicate entries 
+(8 have 2 entries and 4 have 3 entries), the only difference between their entries being in the 
+"Virtual Services Offered" section. 
 The entire Yelp dataset is huge and much of it is not necessary for my domain of bars.

### 2021-02-03 filter.py

@@ -1,24 +1,69 @@
 
+import os
 import json
 import pandas as pd
 
 
-business_json_path = 'yelp_dataset/yelp_academic_dataset_business.json'
-'''
-totalCount = 0
-openCount = 0
-with open('yelp_dataset/yelp_academic_dataset_business.json','r') as f:
-    with open('processedData/business_filtered.json','w') as fout:
-        for line in f:
+def getIDs():
+    inFile = open('dataset/business.json','r')
+    outFile = open('processed/business_ids.json','w')
+    ids = {}
+    count = 0
+    for line in inFile:
+        business = json.loads(line)
+        ID = business['business_id']
+        try:
+            if business['is_open'] == 1 and 'Restaurant' in business['categories']:
+                ids[ID] = 1
+        except:
+            pass
+    json.dump(ids,outFile)
+    inFile.close()
+    outFile.close()
+    return ids
+
+def filterRelevant():
+    files = ['business.json','checkin.json','review.json','tip.json','covid_features.json']
+    for fileName in files:
+        inFile = open('dataset/'+fileName,'r')
+        outFile = open('processed/'+fileName,'w')
+        for line in inFile:
             business = json.loads(line)
-            isOpen = business['is_open']
-            if isOpen == 1:
-                openCount += 1
-                fout.write(line)
-            totalCount += 1
-print(openCount)
-print(totalCount)
+            if business['business_id'] in ids:
+                outFile.write(line)
+        inFile.close()
+        outFile.close()
+
+def toCSV():
+    files = ['business','checkin','review','tip','covid_features']
+    for fileName in files:
+        panda = pd.read_json('processed/'+fileName+'.json', lines=True)
+        panda.to_csv('processed/'+fileName+'.csv',index=False)
+
+#ids = getIDs()
+#filterRelevant()
+#toCSV()
+
+inFile = open('processed/covid_features.json')
+duplicates = ['D35unF350QG06LXDe0Cv-Q',
+'nLGgJBtMfGv7_juXipXheA',
+'b5uCtk5iUcdeSWBNAx40Ug',
+'l3joBBpkq0ib11dKUpKMAw',
+'4v0aj1VWtNwvzOeQ2Ohzrg',
+'BpL-s3p572cxNdXJAsahpA',
+'ZBfoUKOkfocsYK6TFQpwaA',
+'s2Fv9gzUnnpfnkUTAMSK7w']
+for line in inFile:
+    business = json.loads(line)
[…]

### 2021-02-03 code.py
made user_ids json

@@ -25,6 +25,7 @@
 
 # bars on yelp
 
+import json
 
 def collaborativeFiltering():
     pass
@@ -103,9 +104,23 @@ Which city are you closest to?
             pass
     return cities[choice-1]
 
+def getName():
+    count = 0
+    name = ''
+    while name == '':
+        name = input('What is your name? > ').strip()
+    users = open('dataset/user.json','r')
+    for line in users:
+        user = json.loads(line)
+        if user['name'] == name:
+            #print(user['user_id'])
+            count += 1
+    print(count)
+
 
 if __name__ == '__main__':
     welcome()
+    getName()
     city = whichCity()
     print(city)
     

### 2021-02-03 filter.py

@@ -1,5 +1,4 @@
 
-import os
 import json
 import pandas as pd
 
@@ -8,7 +7,6 @@ def getIDs():
     inFile = open('dataset/business.json','r')
     outFile = open('processed/business_ids.json','w')
     ids = {}
-    count = 0
     for line in inFile:
         business = json.loads(line)
         ID = business['business_id']
@@ -40,25 +38,33 @@ def toCSV():
         panda = pd.read_json('processed/'+fileName+'.json', lines=True)
         panda.to_csv('processed/'+fileName+'.csv',index=False)
 
+def covidDuplicates():
+    inFile = open('processed/covid_features.json')
+    duplicates = ['D35unF350QG06LXDe0Cv-Q','nLGgJBtMfGv7_juXipXheA','b5uCtk5iUcdeSWBNAx40Ug','l3joBBpkq0ib11dKUpKMAw','4v0aj1VWtNwvzOeQ2Ohzrg','BpL-s3p572cxNdXJAsahpA','ZBfoUKOkfocsYK6TFQpwaA','s2Fv9gzUnnpfnkUTAMSK7w']
+    for line in inFile:
+        business = json.loads(line)
+        ID = business['business_id']
+        if ID in duplicates:
+            print(line)
+
+def getUserIDs():
+    inFile = open('dataset/user.json','r')
+    outFile = open('processed/user_ids.json','w')
+    userIDs = {}
+    for line in inFile:
+        user = json.loads(line)
+        userIDs[user['user_id']] = user['name']
+    json.dump(userIDs,outFile)
+    inFile.close()
+    outFile.close()
+    return userIDs
+
+
 #ids = getIDs()
 #filterRelevant()
 #toCSV()
-
-inFile = open('processed/covid_features.json')
-duplicates = ['D35unF350QG06LXDe0Cv-Q',
-'nLGgJBtMfGv7_juXipXheA',
-'b5uCtk5iUcdeSWBNAx40Ug',
-'l3joBBpkq0ib11dKUpKMAw',
-'4v0aj1VWtNwvzOeQ2Ohzrg',
-'BpL-s3p572cxNdXJAsahpA',
-'ZBfoUKOkfocsYK6TFQpwaA',
-'s2Fv9gzUnnpfnkUTAMSK7w']
-for line in inFile:
-    business = json.loads(line)
-    ID = business['business_id']
-    if ID in duplicates:
-        print(line)
-
+#covidDuplicates()
+userIDs = getUserIDs()
 
[…]

### 2021-02-04 paper/paper.tex
allowing user ID entry

@@
 "Virtual Services Offered" section. 
+Allow to pick user ID of 10 users with most reviews made, to avoid cold start problem
+4 users have over 10,000 reviews, the most is 14,455 (Victor, 8k3aO-mPeyhbR5HUucA5aA). 
+I actually want to select the most reviews of the places I am accepting, restaurants. 
+When taking the most reviews of restaurants in teh dataset, a threshold of 600 reviews gives a list of 10 users, 
+with the most being 1747. 
+I then checked each of these for how many different restaurants they had reviewed and the greatest number of times 
+they had reviewed a single restaurant to check that none of them were just users who reviewed a small number of 
+restaurants a huge amount of times each. 
+All 10 users passed this criteria, with even the least active user having reviewed 496 unique restaurants, which will 
+certainly avoid any cold-start problems. 
 The entire Yelp dataset is huge and much of it is not necessary for my domain of bars.

### 2021-02-04 code.py

@@ -106,24 +106,24 @@ Which city are you closest to?
 
 def getName():
     count = 0
-    name = ''
-    while name == '':
-        name = input('What is your name? > ').strip()
-    users = open('dataset/user.json','r')
-    for line in users:
-        user = json.loads(line)
-        if user['name'] == name:
-            #print(user['user_id'])
-            count += 1
-    print(count)
+    ID = ''
+    users = json.load(open('processed/user_ids.json','r'))
+    while ID == '':
+        ID = input('Enter your user ID > ').strip()
+        try:
+            name = users[ID]
+        except:
+            ID = ''
+        print('Hello, {}'.format(name))
+        ID = ''
 
 
 if __name__ == '__main__':
     welcome()
     getName()
-    city = whichCity()
-    print(city)
-    
+    #city = whichCity()
+    #print(city)
+    '''
     while True:
         choice = menu()
         if choice == 1:
@@ -132,6 +132,7 @@ if __name__ == '__main__':
             pass
         elif choice == 3:
             info()
+    '''
 
 # Resources:
 '''

### 2021-02-04 filter.py

@@ -59,41 +59,87 @@ def getUserIDs():
     outFile.close()
     return userIDs
 
+def mostActiveUsers():
+    inFile = open('dataset/user.json','r')
+    #outFile = open('processed/activeUsers.json','w')
+    userIDs = {}
+    for line in inFile:
+        user = json.loads(line)
+        count = user['review_count']
+        if count > 10000:
+            print(count)
+    #json.dump(userIDs,outFile)
+    inFile.close()
+    #outFile.close()
+    #return userIDs
+
+def numberRestaurantReviews():
+    inFile = open('processed/review.json','r')
+    outFile = open('processed/numberReviews.json','w')
+    userIDs = {}
+    for line in inFile:
+        review = json.loads(line)
+        user = review['user_id']
+        if user in userIDs:
+            userIDs[user] += 1
+        else:
+            userIDs[user] = 1
+    json.dump(userIDs,outFile)
+    inFile.close()
+    outFile.close()
+
+def mostReviews():
+    inFile = open('processed/numberReviews.json','r')
+    userIDs = json.load(inFile)
+    for user in userIDs:
+        if userIDs[user] > 600:
+            print(user, userIDs[user])
+
+users = ['CxDOIDnH8gp9KXzpBHJYXw','ELcQDlf69kb-ihJfxZyL0A','bLbSNkLggFnqwNNzzq-Ijw','U4INQZOPSUaj8hMjLlZ3KA',
+'DK57YibC5ShBmqQl97CKog','d_TBs6J3twMy9GChqUEXkg','PKEzKWv_FktMm2mGPjwd0Q','cMEtAiW60I5wE_vLfTxoJQ',
+'MMf0LhEk5tGa1LvN7zcDnA','V-BbqKqO8anwplGRx9Q5aQ']
+
+def filterMostActiveReviews():
+    inFile = open('processed/review.json','r')
+    outFile = open('processed/mostActiveReviews.json','w')
+    for line in inFile:
+        review = json.loads(line)
+        if review['user_id'] in users:
+            outFile.write(line)
+    inFile.close()
+    outFile.close()
+
+def uniqueRestaurants():
+    for ID in users:
+        inFile = open('processed/mostActiveReviews.json','r')
+        businesses = {}
+        for line in inFile:
+            review = json.loads(line)
+            if review['user_id'] == ID:
+                business = review['business_id']
+                if business in businesses:
+                    businesses[business] += 1
+                else:
+                    businesses[business] = 1
[…]

### 2021-02-05 paper/paper.tex
going to filter restaurants by nearest city

@@
 The input interface is the command line.
-The program offers users opportunites to input information about themselves,
+The program offers users opportunities to input information about themselves,
 as well as make choices from a selection of items which the system provides or suggests.

### 2021-02-05 code.py

@@ -23,15 +23,57 @@
 #   limitations
 # in order to get an idea of common or good or useful combinations of systems
 
-# bars on yelp
+# restaurants on yelp
 
+import os
 import json
+from collections import defaultdict
+#from surprise import BaselineOnly
+from surprise import SVD
+from surprise import Dataset
+from surprise import Reader
+#from surprise.model_selection import cross_validate
+
+
+def get_top_n(predictions, n=10):
+    top_n = defaultdict(list)
+    for uid, iid, true_r, est, _ in predictions:
+        top_n[uid].append((iid, est))
+    for uid, user_ratings in top_n.items():
+        user_ratings.sort(key=lambda x: x[1], reverse=True)
+        top_n[uid] = user_ratings[:n]
+    return top_n
+
+def getRecommendations(uid):
+    file_path = os.path.expanduser('processed/mostActiveReviews.csv')
+    reader = Reader(line_format='user item rating timestamp', sep=',')
+    data = Dataset.load_from_file(file_path, reader=reader)
+    #cross_validate(BaselineOnly(), data, verbose=True)
+
+    trainset = data.build_full_trainset()
+    algo = SVD()
+    algo.fit(trainset)
+
+    testset = trainset.build_anti_testset()
+    predictions = algo.test(testset)
+
+    top_n = get_top_n(predictions, n=10)
+
+    businessesFile = open('processed/business_ids.json','r')
+    businesses = json.load(businessesFile)
+    businessesFile.close()
+    usersFile = open('processed/user_ids.json','r')
+    users = json.load(usersFile)
+    usersFile.close()
+    #print(top_n.items())
+    for userid, user_ratings in top_n.items():
+        if userid == uid:
+            print(users[uid])
+            for (iid, _) in user_ratings:
+                print(businesses[iid])
+            #print(users[uid], [businesses[iid] for (iid, _) in user_ratings])
 
 def collaborativeFiltering():
-    pass
-
-
-def getRecommendations():
     pass
 
 def getPreferences():
@@ -112,15 +154,16 @@ def getName():
         ID = input('Enter your user ID > ').strip()
         try:
             name = users[ID]
+            print('Hello, {}'.format(name))
         except:
             ID = ''
-        print('Hello, {}'.format(name))
-        ID = ''
+    return ID
 
 
 if __name__ == '__main__':
     welcome()
-    getName()
+    uid = getName()
+    getRecommendations(uid)
     #city = whichCity()
     #print(city)
     '''

### 2021-02-05 filter.py

@@ -12,7 +12,7 @@ def getIDs():
         ID = business['business_id']
         try:
             if business['is_open'] == 1 and 'Restaurant' in business['categories']:
-                ids[ID] = 1
+                ids[ID] = business['name']
         except:
             pass
     json.dump(ids,outFile)
@@ -95,11 +95,10 @@ def mostReviews():
         if userIDs[user] > 600:
             print(user, userIDs[user])
 
-users = ['CxDOIDnH8gp9KXzpBHJYXw','ELcQDlf69kb-ihJfxZyL0A','bLbSNkLggFnqwNNzzq-Ijw','U4INQZOPSUaj8hMjLlZ3KA',
-'DK57YibC5ShBmqQl97CKog','d_TBs6J3twMy9GChqUEXkg','PKEzKWv_FktMm2mGPjwd0Q','cMEtAiW60I5wE_vLfTxoJQ',
-'MMf0LhEk5tGa1LvN7zcDnA','V-BbqKqO8anwplGRx9Q5aQ']
-
 def filterMostActiveReviews():
+    users = ['CxDOIDnH8gp9KXzpBHJYXw','ELcQDlf69kb-ihJfxZyL0A','bLbSNkLggFnqwNNzzq-Ijw','U4INQZOPSUaj8hMjLlZ3KA',
+    'DK57YibC5ShBmqQl97CKog','d_TBs6J3twMy9GChqUEXkg','PKEzKWv_FktMm2mGPjwd0Q','cMEtAiW60I5wE_vLfTxoJQ',
+    'MMf0LhEk5tGa1LvN7zcDnA','V-BbqKqO8anwplGRx9Q5aQ']
     inFile = open('processed/review.json','r')
     outFile = open('processed/mostActiveReviews.json','w')
     for line in inFile:
@@ -124,6 +123,20 @@ def uniqueRestaurants():
         inFile.close()
         print(len(businesses),businesses[max(businesses,key=businesses.get)])
 
+def dropText():
+    panda = pd.read_json('processed/mostActiveReviews.json',lines=True)
+    panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
+    panda.to_csv('processed/mostActiveReviews.csv',index=False)
+
+
+def splitCities():
+    cities = {'Montreal':(45.50884,-73.58781),'Calgary':(51.05011,-114.08529),'Toronto':(43.70011,-79.4163),
[…]

### 2021-02-06 paper/paper.tex
split restaurants by closest city

@@
 certainly avoid any cold-start problems. 
+I split the businesses into 10 groups, by the city which they are closest to, I did this by using latitude and 
+longitude rather than State or 
 The entire Yelp dataset is huge and much of it is not necessary for my domain of bars.

### 2021-02-06 code.py

@@ -60,12 +60,13 @@ def getRecommendations(uid):
     top_n = get_top_n(predictions, n=10)
 
     businessesFile = open('processed/business_ids.json','r')
-    businesses = json.load(businessesFile)
-    businessesFile.close()
-    usersFile = open('processed/user_ids.json','r')
-    users = json.load(usersFile)
-    usersFile.close()
-    #print(top_n.items())
+    #businesses = json.load(businessesFile)
+    #businessesFile.close()
+    #usersFile = open('processed/user_ids.json','r')
+    #users = json.load(usersFile)
+    #usersFile.close()
+    print(top_n.items())
+    return top_n.items()
     for userid, user_ratings in top_n.items():
         if userid == uid:
             print(users[uid])

### 2021-02-06 filter.py

@@ -134,8 +134,42 @@ def splitCities():
     'Pittsburgh':(40.44062,-79.99589),'Charlotte':(35.22709,-80.84313),'Urbana-Champaign':(40.11059,-88.20727),
     'Phoenix':(33.44838,-112.07404),'Las Vegas':(36.17497,-115.13722),'Madison':(43.07305,-89.40123),
     'Cleveland':(41.4995,-81.69541)}
-    inFile = open
-    print(len(cities))
+    inFile = open('processed/business.json','r')
+    outFile = open('processed/closestCity.json','w')
+    closest = {}
+    for line in inFile:
+        business = json.loads(line)
+        lat = business['latitude']
+        lon = business['longitude']
+        smallest = 100000
+        closestCity = ''
+        for city in cities:
+            coords = cities[city]
+            distance = (coords[0]-lat)**2+(coords[1]-lon)**2
+            if distance < smallest:
+                smallest = distance
+                closestCity = city
+        closest[business['business_id']] = closestCity
+        with open('processed/cities/'+closestCity+'.json','a') as fout:
+            fout.write(line)
+    json.dump(closest,outFile)
+    inFile.close()
+    outFile.close()
+
+def splitIDs():
+    cities = ['Montreal','Calgary','Toronto','Pittsburgh','Charlotte',
+    'Urbana-Champaign','Phoenix','Las Vegas','Madison','Cleveland']
+    for city in cities:
+        ids = {}
+        inFile = open('processed/cities/'+city+'.json','r')
+        outFile = open('processed/cities/'+city+'_ids.json','w')
+        for line in inFile:
+            business = json.loads(line)
+            ids[business['business_id']] = business['name']
+        json.dump(ids,outFile)
+        inFile.close()
+        outFile.close()
+    
 
 #ids = getIDs()
 #filterRelevant()
@@ -148,7 +182,8 @@ def splitCities():
 #filterMostActiveReviews()
 #uniqueRestaurants()
 #dropText()
-splitCities()
+#splitCities()
+splitIDs()
 
 # user features: user_id, name, review_count, yelping_since, useful, funny, cool, elite, friends, fans, average_stars, compliment_hot, compliment_more, compliment_profile, compliment_cute, compliment_list, compliment_note, compliment_plain, compliment_cool, compliment_funny, compliment_writer, compliment_photos
 

### 2021-02-07 code.py
tried using SVD on all reviews, didn't work

@@ -33,6 +33,7 @@ from surprise import SVD
 from surprise import Dataset
 from surprise import Reader
 #from surprise.model_selection import cross_validate
+import texttable
 
 
 def get_top_n(predictions, n=10):
@@ -58,21 +59,35 @@ def getRecommendations(uid):
     predictions = algo.test(testset)
 
     top_n = get_top_n(predictions, n=10)
-
-    businessesFile = open('processed/business_ids.json','r')
-    #businesses = json.load(businessesFile)
-    #businessesFile.close()
-    #usersFile = open('processed/user_ids.json','r')
-    #users = json.load(usersFile)
-    #usersFile.close()
-    print(top_n.items())
+    end = time.time()
+    #print(top_n.items())
     return top_n.items()
-    for userid, user_ratings in top_n.items():
+
+def presentRecommendations(uid,items):
+    businessesFile = open('processed/business_ids.json','r')
+    businesses = json.load(businessesFile)
+    businessesFile.close()
+    usersFile = open('processed/user_ids.json','r')
+    users = json.load(usersFile)
+    usersFile.close()
+    table = texttable.Texttable()
+    table.set_cols_dtype(['i','t','t','t'])
+    table.set_cols_align(['c','l','c','c'])
+    table.set_cols_valign(['m','m','m','m'])
+    rows = [['Number','Name', 'City', 'Stars']]
+    for userid, user_ratings in items:
         if userid == uid:
-            print(users[uid])
+            #print(users[uid])
+            i = 1
             for (iid, _) in user_ratings:
-                print(businesses[iid])
+                business = businesses[iid]
+                rows.append([i,business['name'],business['city'],business['stars']])
+                i += 1
+                #print(businesses[iid])
             #print(users[uid], [businesses[iid] for (iid, _) in user_ratings])
+    table.add_rows(rows)
+    print(table.draw() + "\n")
+
 
 def collaborativeFiltering():
     pass
@@ -164,7 +179,8 @@ def getName():
 if __name__ == '__main__':
     welcome()
     uid = getName()
-    getRecommendations(uid)
+    items = getRecommendations(uid)
+    presentRecommendations(uid,items)
     #city = whichCity()
     #print(city)
     '''

### 2021-02-07 filter.py

@@ -12,7 +12,9 @@ def getIDs():
         ID = business['business_id']
         try:
             if business['is_open'] == 1 and 'Restaurant' in business['categories']:
-                ids[ID] = business['name']
+                ids[ID] = {}
+                for attribute in ['name','address','city','state','postal_code','latitude','longitude','stars','review_count','attributes','categories','hours']:
+                    ids[ID][attribute] = business[attribute]
         except:
             pass
     json.dump(ids,outFile)
@@ -124,10 +126,20 @@ def uniqueRestaurants():
         print(len(businesses),businesses[max(businesses,key=businesses.get)])
 
 def dropText():
-    panda = pd.read_json('processed/mostActiveReviews.json',lines=True)
-    panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
-    panda.to_csv('processed/mostActiveReviews.csv',index=False)
-
+    #panda = pd.read_json('processed/mostActiveReviews.json',lines=True)
+    #panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
+    #panda.to_csv('processed/mostActiveReviews.csv',index=False)
+    inFile = open('processed/review.json','r')
+    outFile = open('processed/review.csv','w')
+    for line in inFile:
+        review = json.loads(line)
+        #outFile.write(review['user_id']+','+review['business_id']+','+str(review['stars'])+','+review['date']+'\n')
+        outFile.write('{},{},{},{}\n'.format(review['user_id'],review['business_id'],int(review['stars']),review['date']))
+    inFile.close()
+    outFile.close()
+    #panda = pd.read_json('processed/review.json',lines=True)
+    #panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
+    #panda.to_csv('processed/review.csv',index=False)
 
 def splitCities():
     cities = {'Montreal':(45.50884,-73.58781),'Calgary':(51.05011,-114.08529),'Toronto':(43.70011,-79.4163),
@@ -181,9 +193,9 @@ def splitIDs():
 #mostReviews()
 #filterMostActiveReviews()
 #uniqueRestaurants()
-#dropText()
+dropText()
 #splitCities()
-splitIDs()
+#splitIDs()
 
 # user features: user_id, name, review_count, yelping_since, useful, funny, cool, elite, friends, fans, average_stars, compliment_hot, compliment_more, compliment_profile, compliment_cute, compliment_list, compliment_note, compliment_plain, compliment_cool, compliment_funny, compliment_writer, compliment_photos
 

### 2021-02-07 paper/paper.tex
tried using SVD on all reviews, didn't work
about to add item-based CF

--- 2021-02-05 paper/paper.tex
@@
 certainly avoid any cold-start problems. 
+I split the businesses into 10 groups, by the city which they are closest to, I did this by using latitude and 
+longitude rather than State or City since this does not give actual distance and restaurants close to a border 
+could then be mis-categorised. 
 The entire Yelp dataset is huge and much of it is not necessary for my domain of bars.
@@
 The first recommender system is collaborative filtering.
+I am using item-based collaborative filtering, since the number of users is larger than the number of items 
+(1,968,703 users vs 209,393 businesses in the entire dataset), this gives greater accuracy of predictions. 
+In this case the list of available items does not change, so a user-based method would be unnecessary. 
+One of the requirements for this system is justifiability, and item-based methods make it much easier to 
+explain why certain predictions were made. 
 I am using a weighted mixed combination of these two systems in order to produce the results

### 2021-02-07 code.py

@@ -37,12 +37,16 @@ import texttable
 
 
 def get_top_n(predictions, n=10):
+    inFile = open('processed/closestCity.json','r')
+    closestCity = json.load(inFile)
+    inFile.close()
     top_n = defaultdict(list)
-    for uid, iid, true_r, est, _ in predictions:
-        top_n[uid].append((iid, est))
-    for uid, user_ratings in top_n.items():
+    for uidLocal, iid, true_r, est, _ in predictions:
+        if uidLocal == uid and closestCity[iid] == city: # only get recommendations for active user and in city they specified
+            top_n[uidLocal].append((iid, est))
+    for uidLocal, user_ratings in top_n.items():
         user_ratings.sort(key=lambda x: x[1], reverse=True)
-        top_n[uid] = user_ratings[:n]
+        top_n[uidLocal] = user_ratings[:n]
     return top_n
 
 def getRecommendations(uid):
@@ -58,12 +62,11 @@ def getRecommendations(uid):
     testset = trainset.build_anti_testset()
     predictions = algo.test(testset)
 
-    top_n = get_top_n(predictions, n=10)
-    end = time.time()
+    top_n = get_top_n(predictions, n=8) # get top 8 recommendations
     #print(top_n.items())
     return top_n.items()
 
-def presentRecommendations(uid,items):
+def presentRecommendations(uid,items): # takes items from recommender and 
     businessesFile = open('processed/business_ids.json','r')
     businesses = json.load(businessesFile)
     businessesFile.close()
@@ -71,10 +74,10 @@ def presentRecommendations(uid,items):
     users = json.load(usersFile)
     usersFile.close()
     table = texttable.Texttable()
-    table.set_cols_dtype(['i','t','t','t'])
-    table.set_cols_align(['c','l','c','c'])
-    table.set_cols_valign(['m','m','m','m'])
-    rows = [['Number','Name', 'City', 'Stars']]
+    table.set_cols_dtype(['i','t','t','t']) # specify data types
+    table.set_cols_align(['c','l','c','c']) # align columns horizontally
+    table.set_cols_valign(['m','m','m','m']) # align columns vertically
+    rows = [['Number','Name', 'City', 'Stars']] # titles of columns
     for userid, user_ratings in items:
         if userid == uid:
             #print(users[uid])
@@ -144,8 +147,8 @@ def welcome():
 def whichCity():
[…]

### 2021-02-08 code.py
implemented item-based CF but takes ages

@@ -34,7 +34,106 @@ from surprise import Dataset
 from surprise import Reader
 #from surprise.model_selection import cross_validate
 import texttable
+import pandas as pd
+import numpy as np
+import math
+import pickle
 
+
+def collaborativeFiltering():
+    ratings = pd.read_csv('processed/mostActiveReviewsCF.csv', encoding='"ISO-8859-1"')
+    ratings = ratings.sample(frac=0.01)
+    ratings_training = ratings.sample(frac=0.7)
+    ratings_test = ratings.drop(ratings_training.index)
+    rating_mean = ratings_training.groupby(['business_id'], as_index=False, sort=False).mean().rename(columns={'rating':'rating_mean'})[['business_id','rating_mean']]
+    adjusted_ratings = pd.merge(ratings_training,rating_mean,on='business_id',how='left',sort=False)
+    adjusted_ratings['rating_adjusted']=adjusted_ratings['rating']-adjusted_ratings['rating_mean']
+    adjusted_ratings.loc[adjusted_ratings['rating_adjusted']==0,'rating_adjusted'] = 1e-8
+    w_matrix = build_w_matrix(adjusted_ratings,True)
+    #predict(uid,iid,w_matrix,adjusted_ratings,rating_mean)
+    recommended_restaurants = recommend(uid,w_matrix,adjusted_ratings,rating_mean)
+    print(recommended_restaurants)
+
+def build_w_matrix(adjusted_ratings, load_existing_w_matrix):
+    w_matrix_columns = ['business_1', 'business_2', 'weight']
+    w_matrix = pd.DataFrame(columns=w_matrix_columns)
+    if load_existing_w_matrix:
+        with open('processed/w_matrix.pkl','rb') as input:
+            w_matrix = pickle.load(input)
+        input.close()
+    else:
+        distinct_business = np.unique(adjusted_ratings['business_id'])
+        i=0
+        for business_1 in distinct_business:
+            user_data = adjusted_ratings[adjusted_ratings['business_id'] == business_1]
+            distinct_users = np.unique(user_data['user_id'])
+
+            record_row_columns = ['user_id','business_1','business_2','rating_adjusted_1','rating_adjusted_2']
+            record_business_1_2 = pd.DataFrame(columns=record_row_columns)
+
+            for c_user_id in distinct_users:
+                c_business_1_rating = user_data[user_data['user_id'] == c_user_id]['rating_adjusted'].iloc[0]
[…]

### 2021-02-08 filter.py

@@ -129,6 +129,10 @@ def dropText():
     #panda = pd.read_json('processed/mostActiveReviews.json',lines=True)
     #panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
     #panda.to_csv('processed/mostActiveReviews.csv',index=False)
+    panda = pd.read_json('processed/mostActiveReviews.json',lines=True)
+    panda = panda.drop(['review_id','useful','funny','cool','text','date'],axis=1)
+    panda.to_csv('processed/mostActiveReviewsCF.csv',index=False)
+    '''
     inFile = open('processed/review.json','r')
     outFile = open('processed/review.csv','w')
     for line in inFile:
@@ -137,9 +141,7 @@ def dropText():
         outFile.write('{},{},{},{}\n'.format(review['user_id'],review['business_id'],int(review['stars']),review['date']))
     inFile.close()
     outFile.close()
-    #panda = pd.read_json('processed/review.json',lines=True)
-    #panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
-    #panda.to_csv('processed/review.csv',index=False)
+    '''
 
 def splitCities():
     cities = {'Montreal':(45.50884,-73.58781),'Calgary':(51.05011,-114.08529),'Toronto':(43.70011,-79.4163),
@@ -181,7 +183,19 @@ def splitIDs():
         json.dump(ids,outFile)
         inFile.close()
         outFile.close()
-    
+
+def splitReviews():
+    inFile = open('processed/review.json','r')
+    closestCity = open('processed/closestCity.json','r')
+    closest = json.load(closestCity)
+    closestCity.close()
+    for line in inFile:
+        business = json.loads(line)
+        closestCity = closest[business['business_id']]
+        with open('processed/cities/'+closestCity+'_reviews.json','a') as fout:
+            fout.write(line)
+    inFile.close()
+
 
 #ids = getIDs()
 #filterRelevant()
@@ -193,50 +207,12 @@ def splitIDs():
 #mostReviews()
 #filterMostActiveReviews()
 #uniqueRestaurants()
-dropText()
+#dropText()
 #splitCities()
 #splitIDs()
+splitReviews()
 
[…]

### 2021-02-08 paper/paper.tex
implemented item-based CF but takes ages
filtering restaurants into 12 categories

@@
 Of the 209,393 businesses in the dataset, 168,903 are still open. 
-There are 43,980 restaurants which are open and contain 'Restaurant' as one of the categories. 
-502 businesses in the overall dataset don't list anything in their categories section. 
-43,052 businesses had at least 1 check-in. 
-There are 4,223,201 reviews of restaurants. 
-695,951 tips for these restaurants. 
+There are 43,965 restaurants which are open and contain 'Restaurants' as one of the categories. 
+I then picked the top 12 most popular categories of restaurant: American, Mexican, Italian, Chinese, 
+Seafood, Japanese, Canadian, Mediterranean, Indian, Thai, Middle Eastern, Vietnamese (in that order). 
+When filtered for restaurants which fall into at least one of these categories, 
+20,565 had at least one check-in in 2019
+There are 2,956,316 reviews of these restaurants. 
+481,073 tips for these restaurants. 
 1,968,703 users. 
-43,992 restaurants listed with covid features, 8 restaurants which make up 12 duplicate entries 
-(8 have 2 entries and 4 have 3 entries), the only difference between their entries being in the 
+None of these restaurants are temporarily closed according to the covid features data. 
+20,573 restaurants listed with covid features, 4 restaurants which make up 8 duplicate entries 
+(4 have 3 entries), the only difference between their entries being in the 
 "Virtual Services Offered" section. 
@@
 could then be mis-categorised. 
+Since there are 4,223,201 reviews of restaurants I filtered them by removing any which had zero 'useful' votes. 
+When filtered to reviews with at least 1 'useful' vote, there are 1,735,732 and when at least 2 votes, 836,929
+At least 3 votes, 462,412, at least 5 votes 190,902. 
+51,737 with at least 10 votes
+
+Top categories: American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, Mediterranean, Indian, Thai, Middle Eastern, Vietnamese
+
 The entire Yelp dataset is huge and much of it is not necessary for my domain of bars.
@@
 Meaning better recommendations than either algorithm could achieve individually.
+Meta-level HRS with collaborative and content-based, or 
+Cascade with knowledge-based and collaborative
+Feature Augmentation with Knowledge Based and Collaborative
 
@@
 explain why certain predictions were made. 
+Item-item collaborative filtering looks for restaurants that are similar, in terms of how people rate them, 
+to the restaurants that the user has already rated and recommend the most similar restaurants. 
+Filtering reviews by city and then using CF won't work because the active user may not have ever reviewed a restaurant 
+in that city, so there is no way to assess similar restaurants to ones they have already rated. 
+Instead I carried out CF on all high quality reviews and produce an ordering of recommendations, which I then filter 
+by city.
+
 I am using a weighted mixed combination of these two systems in order to produce the results

### 2021-02-08 code.py

@@ -42,14 +42,15 @@ import pickle
 
 def collaborativeFiltering():
     ratings = pd.read_csv('processed/mostActiveReviewsCF.csv', encoding='"ISO-8859-1"')
-    ratings = ratings.sample(frac=0.01)
+    #ratings = ratings.sample(frac=1)
     ratings_training = ratings.sample(frac=0.7)
     ratings_test = ratings.drop(ratings_training.index)
     rating_mean = ratings_training.groupby(['business_id'], as_index=False, sort=False).mean().rename(columns={'rating':'rating_mean'})[['business_id','rating_mean']]
     adjusted_ratings = pd.merge(ratings_training,rating_mean,on='business_id',how='left',sort=False)
     adjusted_ratings['rating_adjusted']=adjusted_ratings['rating']-adjusted_ratings['rating_mean']
     adjusted_ratings.loc[adjusted_ratings['rating_adjusted']==0,'rating_adjusted'] = 1e-8
-    w_matrix = build_w_matrix(adjusted_ratings,True)
+    w_matrix = build_w_matrix(adjusted_ratings,False)
+    print('built w_matrix')
     #predict(uid,iid,w_matrix,adjusted_ratings,rating_mean)
     recommended_restaurants = recommend(uid,w_matrix,adjusted_ratings,rating_mean)
     print(recommended_restaurants)

### 2021-02-08 filter.py

@@ -4,17 +4,25 @@ import pandas as pd
 
 
 def getIDs():
-    inFile = open('dataset/business.json','r')
+    inFile = open('processed/2019business.json','r')
     outFile = open('processed/business_ids.json','w')
     ids = {}
+    categories = ['American', 'Mexican', 'Italian', 'Chinese', 'Seafood', 'Japanese', 'Canadian', 'Mediterranean', 'Indian', 'Thai', 'Middle Eastern', 'Vietnamese']
     for line in inFile:
         business = json.loads(line)
         ID = business['business_id']
+        busCategories = business['categories']
         try:
-            if business['is_open'] == 1 and 'Restaurant' in business['categories']:
-                ids[ID] = {}
-                for attribute in ['name','address','city','state','postal_code','latitude','longitude','stars','review_count','attributes','categories','hours']:
-                    ids[ID][attribute] = business[attribute]
+            if business['is_open'] == 1 and 'Restaurants' in busCategories:
+                check = False
+                for category in categories:
+                    if category in busCategories:
+                        check = True
+                        break
+                if check:
+                    ids[ID] = {}
+                    for attribute in ['name','address','city','state','postal_code','latitude','longitude','stars','review_count','attributes','categories','hours']:
+                        ids[ID][attribute] = business[attribute]
         except:
             pass
     json.dump(ids,outFile)
@@ -33,6 +41,33 @@ def filterRelevant():
                 outFile.write(line)
         inFile.close()
         outFile.close()
+
+def lastCheckin():
+    inFile = open('processed/checkin.json','r')
+    outFile = open('processed/lastCheckin.json','w')
+    last = {}
+    for line in inFile:
+        business = json.loads(line)
+        ID = business['business_id']
+        checkins = business['date'].split(', ')
+        last[ID] = checkins[-1]
+    json.dump(last,outFile)
+    inFile.close()
+    outFile.close()
+
+def recentCheckin():
+    lastFile = open('processed/lastCheckin.json','r')
+    last = json.load(lastFile)
+    lastFile.close()
+    inFile = open('processed/business.json','r')
+    outFile = open('processed/2019business.json','w')
+    for line in inFile:
[…]

### 2021-02-09 code.py
added asking for category of food

@@ -38,22 +38,27 @@ import pandas as pd
 import numpy as np
 import math
 import pickle
-
+import time
 
 def collaborativeFiltering():
-    ratings = pd.read_csv('processed/mostActiveReviewsCF.csv', encoding='"ISO-8859-1"')
-    #ratings = ratings.sample(frac=1)
+    ratings = pd.read_csv('processed/usefulReviews.csv', encoding='"ISO-8859-1"')
+    ratings = ratings.sample(frac=0.2)
     ratings_training = ratings.sample(frac=0.7)
     ratings_test = ratings.drop(ratings_training.index)
     rating_mean = ratings_training.groupby(['business_id'], as_index=False, sort=False).mean().rename(columns={'rating':'rating_mean'})[['business_id','rating_mean']]
     adjusted_ratings = pd.merge(ratings_training,rating_mean,on='business_id',how='left',sort=False)
     adjusted_ratings['rating_adjusted']=adjusted_ratings['rating']-adjusted_ratings['rating_mean']
     adjusted_ratings.loc[adjusted_ratings['rating_adjusted']==0,'rating_adjusted'] = 1e-8
+    start = time.time()
     w_matrix = build_w_matrix(adjusted_ratings,False)
+    end = time.time()
+    print(end-start)
     print('built w_matrix')
     #predict(uid,iid,w_matrix,adjusted_ratings,rating_mean)
     recommended_restaurants = recommend(uid,w_matrix,adjusted_ratings,rating_mean)
     print(recommended_restaurants)
+    end = time.time()
+    print(end-start)
 
 def build_w_matrix(adjusted_ratings, load_existing_w_matrix):
     w_matrix_columns = ['business_1', 'business_2', 'weight']
@@ -95,12 +100,12 @@ def build_w_matrix(adjusted_ratings, load_existing_w_matrix):
         output.close()
     return w_matrix
 
-def predict(user_id, business_id, w_matrix, adjusted_ratings, rating_mean):
+def predict(uid, business_id, w_matrix, adjusted_ratings, rating_mean):
     if rating_mean[rating_mean['business_id']==business_id].shape[0] > 0:
         mean_rating = rating_mean[rating_mean['business_id']==business_id]['rating_mean'].iloc[0]
     else:
         mean_rating = 2.5
-    user_other_ratings = adjusted_ratings[adjusted_ratings['user_id']==user_id]
+    user_other_ratings = adjusted_ratings[adjusted_ratings['user_id']==uid]
[…]

### 2021-02-09 filter.py

@@ -164,9 +164,9 @@ def dropText():
     #panda = pd.read_json('processed/mostActiveReviews.json',lines=True)
     #panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
     #panda.to_csv('processed/mostActiveReviews.csv',index=False)
-    panda = pd.read_json('processed/reviews10.json',lines=True)
+    panda = pd.read_json('processed/usefulReviews.json',lines=True)
     panda = panda.drop(['review_id','useful','funny','cool','text','date'],axis=1)
-    panda.to_csv('processed/usefulReviewsCF.csv',index=False)
+    panda.to_csv('processed/usefulReviews.csv',index=False)
     '''
     inFile = open('processed/review.json','r')
     outFile = open('processed/review.csv','w')
@@ -220,24 +220,24 @@ def splitIDs():
         outFile.close()
 
 def filterUsefulReviews():
-    inFile = open('processed/reviews15.json','r')
-    outFile = open('processed/reviews20.json','w')
+    inFile = open('processed/review.json','r')
+    outFile = open('processed/usefulReviews.json','w')
     for line in inFile:
         review = json.loads(line)
-        if review['useful'] > 19:
+        if review['useful'] > 9:
             outFile.write(line)
     inFile.close()
     outFile.close()
 
 def splitReviews():
-    inFile = open('processed/reviews20.json','r')
+    inFile = open('processed/usefulReviews.json','r')
     closestCity = open('processed/closestCity.json','r')
     closest = json.load(closestCity)
     closestCity.close()
     for line in inFile:
         review = json.loads(line)
         closestCity = closest[review['business_id']]
-        with open('processed/cities/'+closestCity+'_reviews20.json','a') as fout:
+        with open('processed/cities/'+closestCity+'_usefulReviews.json','a') as fout:
             fout.write(line)
     inFile.close()
 
@@ -269,9 +269,9 @@ def getCategories():
 #mostReviews()
 #filterMostActiveReviews()
 #uniqueRestaurants()
-#dropText()
+dropText()
 #splitCities()
-splitIDs()
+#splitIDs()
 #filterUsefulReviews()
 #splitReviews()
 #getCategories()

### 2021-02-09 code.py
item-based CF is taking too long

@@ -40,9 +40,9 @@ import math
 import pickle
 import time
 
-def collaborativeFiltering():
+def collaborativeFiltering(): # code adapted from https://github.com/wwwbbb8510/baseline-rs
     ratings = pd.read_csv('processed/usefulReviews.csv', encoding='"ISO-8859-1"')
-    ratings = ratings.sample(frac=0.2)
+    ratings = ratings.sample(frac=0.01)
     ratings_training = ratings.sample(frac=0.7)
     ratings_test = ratings.drop(ratings_training.index)
     rating_mean = ratings_training.groupby(['business_id'], as_index=False, sort=False).mean().rename(columns={'rating':'rating_mean'})[['business_id','rating_mean']]
@@ -50,12 +50,19 @@ def collaborativeFiltering():
     adjusted_ratings['rating_adjusted']=adjusted_ratings['rating']-adjusted_ratings['rating_mean']
     adjusted_ratings.loc[adjusted_ratings['rating_adjusted']==0,'rating_adjusted'] = 1e-8
     start = time.time()
-    w_matrix = build_w_matrix(adjusted_ratings,False)
+    w_matrix = build_w_matrix(adjusted_ratings,True)
     end = time.time()
     print(end-start)
     print('built w_matrix')
     #predict(uid,iid,w_matrix,adjusted_ratings,rating_mean)
     recommended_restaurants = recommend(uid,w_matrix,adjusted_ratings,rating_mean)
+    presentRecommendations(recommended_restaurants)
+    #items = []
+    #for i in range(8):
+    #    items.append(recommended_restaurants['business_id'].iloc[i])
+    #print(items)
+    #presentRecommendations(items)
+    
     print(recommended_restaurants)
     end = time.time()
     print(end-start)
@@ -125,7 +132,7 @@ def predict(uid, business_id, w_matrix, adjusted_ratings, rating_mean):
         predicted_rating = mean_rating + sum_weighted_other_ratings/sum_weights
     return predicted_rating
 
-def recommend(uid,w_matrix,adjusted_ratings,rating_mean,amount=8):
+def recommend(uid,w_matrix,adjusted_ratings,rating_mean,amount=5):
     distinct_business = np.unique(adjusted_ratings['business_id'])
     user_ratings_all_business = pd.DataFrame(columns=['business_id','rating'])
     user_rating = adjusted_ratings[adjusted_ratings['user_id']==uid]
[…]

### 2021-02-09 filter.py

@@ -243,17 +243,23 @@ def splitReviews():
 
 def getCategories():
     inFile = open('processed/business.json','r')
+    outFile = open('processed/categories.json','w')
     allCategories = {}
     for line in inFile:
         business = json.loads(line)
-        categories = business['categories'].split(', ')
-        for category in categories:
-            if category in allCategories:
-                allCategories[category] += 1
-            else:
-                allCategories[category] = 1
-    sortedDict = dict(sorted(allCategories.items(),key=lambda item: item[1]))
-    print(sortedDict)
+        #categories = business['categories'].split(', ')
+        allCategories[business['business_id']] = business['categories']
+        #for category in categories:
+        #    allCategories[business]
+        #    if category in allCategories:
+        #        allCategories[category] += 1
+        #    else:
+        #        allCategories[category] = 1
+    #sortedDict = dict(sorted(allCategories.items(),key=lambda item: item[1]))
+    #print(sortedDict)
+    json.dump(allCategories,outFile)
+    inFile.close()
+    outFile.close()
 
 
 
@@ -269,12 +275,12 @@ def getCategories():
 #mostReviews()
 #filterMostActiveReviews()
 #uniqueRestaurants()
-dropText()
+#dropText()
 #splitCities()
 #splitIDs()
 #filterUsefulReviews()
 #splitReviews()
-#getCategories()
+getCategories()
 
 # user features: user_id, name, review_count, yelping_since, useful, funny, cool, elite, friends, fans, average_stars, compliment_hot, compliment_more, compliment_profile, compliment_cute, compliment_list, compliment_note, compliment_plain, compliment_cool, compliment_funny, compliment_writer, compliment_photos
 

### 2021-02-09 paper/paper.tex
implemented SVD and can now see more info about restaurant

@@
 best of both algorithms if done properly. 
+I am using a Cascade Hybrid Recommender System, since this has been shown to accurate predictions 
+in my domain of restaurant recommendations \cite{burke2007hybrid}. 
+The two systems are Collaborative Filtering and Content Based.
+
 Meaning better recommendations than either algorithm could achieve individually.
-Meta-level HRS with collaborative and content-based, or 
 Cascade with knowledge-based and collaborative
-Feature Augmentation with Knowledge Based and Collaborative
 

### 2021-02-09 code.py

@@ -181,50 +181,93 @@ def get_top_n(predictions, n=10):
     return top_n
 
 def getRecommendations(uid):
-    file_path = os.path.expanduser('processed/mostActiveReviews.csv')
+    file_path = os.path.expanduser('processed/usefulReviews.csv')
     reader = Reader(line_format='user item rating timestamp', sep=',')
     data = Dataset.load_from_file(file_path, reader=reader)
     #cross_validate(BaselineOnly(), data, verbose=True)
 
     trainset = data.build_full_trainset()
     algo = SVD()
+    #start = time.time()
+    print('Getting Recommendations...')
     algo.fit(trainset)
+    print()
+    #end = time.time()
+    #print(end-start)
+    suggestions = {}
+    for iid in iids:
+        if iid in categories and category in categories[iid]: # include all reviews?
+            pred = algo.predict(uid,iid)#,verbose=True)#,r_ui=4, verbose=True)
+            suggestions[pred[1]] = pred[3]
+    suggestions = sorted(suggestions.items(), key=lambda item: item[1],reverse=True)
+    #print(suggestions)
+    return suggestions[:8] # present top 8 restaurants
+    #testset = trainset.build_anti_testset()
+    #predictions = algo.test(testset)
 
-    testset = trainset.build_anti_testset()
-    predictions = algo.test(testset)
-
-    top_n = get_top_n(predictions, n=8) # get top 8 recommendations
+    #top_n = get_top_n(predictions, n=8) # get top 8 recommendations
     #print(top_n.items())
-    return top_n.items()
+    #return top_n.items()
 
 def hybrid():
     pass
 
 def presentRecommendations(items): # takes items from recommender and 
-    businessesFile = open('processed/business_ids.json','r')
-    businesses = json.load(businessesFile)
-    businessesFile.close()
-    #usersFile = open('processed/user_ids.json','r')
-    #users = json.load(usersFile)
-    #usersFile.close()
+    if items == []:
+        print('There are no {} restaurants in {}'.format(category,city))
+        return None
     table = texttable.Texttable()
     table.set_cols_dtype(['i','t','t','t']) # specify data types
     table.set_cols_align(['c','l','c','c']) # align columns horizontally
     table.set_cols_valign(['m','m','m','m']) # align columns vertically
     rows = [['Number','Name', 'City', 'Stars']] # titles of columns
     i = 1
-    for iid in items:
-        #for userid, user_ratings in items:
-        #if userid == uid:
-        #    i = 1
-            #for (iid, _) in user_ratings:
[…]

### 2021-02-09 filter.py

@@ -165,7 +165,7 @@ def dropText():
     #panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
     #panda.to_csv('processed/mostActiveReviews.csv',index=False)
     panda = pd.read_json('processed/usefulReviews.json',lines=True)
-    panda = panda.drop(['review_id','useful','funny','cool','text','date'],axis=1)
+    panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
     panda.to_csv('processed/usefulReviews.csv',index=False)
     '''
     inFile = open('processed/review.json','r')
@@ -224,7 +224,7 @@ def filterUsefulReviews():
     outFile = open('processed/usefulReviews.json','w')
     for line in inFile:
         review = json.loads(line)
-        if review['useful'] > 9:
+        if review['useful'] > 4:
             outFile.write(line)
     inFile.close()
     outFile.close()
@@ -275,12 +275,12 @@ def getCategories():
 #mostReviews()
 #filterMostActiveReviews()
 #uniqueRestaurants()
-#dropText()
+dropText()
 #splitCities()
 #splitIDs()
 #filterUsefulReviews()
 #splitReviews()
-getCategories()
+#getCategories()
 
 # user features: user_id, name, review_count, yelping_since, useful, funny, cool, elite, friends, fans, average_stars, compliment_hot, compliment_more, compliment_profile, compliment_cute, compliment_list, compliment_note, compliment_plain, compliment_cool, compliment_funny, compliment_writer, compliment_photos
 

### 2021-02-09 code.py
added covid information to moreInfo

@@ -189,9 +189,8 @@ def getRecommendations(uid):
     trainset = data.build_full_trainset()
     algo = SVD()
     #start = time.time()
-    print('Getting Recommendations...')
+    print('Getting recommendations...')
     algo.fit(trainset)
-    print()
     #end = time.time()
     #print(end-start)
     suggestions = {}
@@ -211,6 +210,7 @@ def getRecommendations(uid):
 
 def hybrid():
     pass
+    # use namsor to get info about person from name, then use to recommend 
 
 def presentRecommendations(items): # takes items from recommender and 
     if items == []:
@@ -228,15 +228,15 @@ def presentRecommendations(items): # takes items from recommender and
         rows.append([i,business['name'],business['city'],business['stars']])
         i += 1
     table.add_rows(rows)
+    print()
     print('Here are some {} restaurants in {} that you might like:'.format(category,city))
     print(table.draw() + '\n')
-    print('Select a restaurant to see more about it')
-    print()
+    print('Select a restaurant to see more information')
     check = False
     while not check:
-        choice = input('Your Choice > ')
+        choice = input('Your Choice > ').strip()
         try:
-            choice = int(choice.strip())
+            choice = int(choice)
             if 1 <= choice <= len(items):
                 check = True
         except:
@@ -247,13 +247,21 @@ def moreInformation(restaurant):
     if restaurant == None:
         return
     business = businesses[restaurant]
+    name = business['name']
     address = '{}\n{}, {}'.format(business['address'],business['city'],business['state'])
+    features = covidFeatures[restaurant]
+    delivery = features['delivery or takeout']
+    delivery = 'Yes' if delivery=='TRUE' else 'No'
+    grubhub = features['Grubhub enabled']
+    grubhub = 'Yes' if grubhub=='TRUE' else 'No'
+    message = features['Covid Banner']
+    closed = features['Temporary Closed Until']
     table = texttable.Texttable()
-    table.set_cols_dtype(['t','t','t','t'])
-    table.set_cols_align(['l','c','c','c'])
-    table.set_cols_valign(['m','m','m','m'])
-    table.add_rows([['Name','Address','Stars','No. reviews'],
-    [business['name'],address,business['stars'],business['review_count']]])
+    table.set_cols_dtype(['t','t','t','t','t','t'])
+    table.set_cols_align(['l','c','c','c','c','c'])
+    table.set_cols_valign(['m','m','m','m','m','m'])
[…]

### 2021-02-09 filter.py

@@ -261,6 +261,20 @@ def getCategories():
     inFile.close()
     outFile.close()
 
+def formatCovidFeatures():
+    inFile = open('processed/covid_features.json','r')
+    outFile = open('processed/covid.json','w')
+    features = {}
+    for line in inFile:
+        business = json.loads(line)
+        ID = business['business_id']
+        features[ID] = {}
+        for attribute in ['highlights','delivery or takeout','Grubhub enabled','Covid Banner','Temporary Closed Until']:
+            features[ID][attribute] = business[attribute]
+    json.dump(features,outFile)
+    inFile.close()
+    outFile.close()
+
 
 
 #ids = getIDs()
@@ -275,12 +289,13 @@ def getCategories():
 #mostReviews()
 #filterMostActiveReviews()
 #uniqueRestaurants()
-dropText()
+#dropText()
 #splitCities()
 #splitIDs()
 #filterUsefulReviews()
 #splitReviews()
 #getCategories()
+formatCovidFeatures()
 
 # user features: user_id, name, review_count, yelping_since, useful, funny, cool, elite, friends, fans, average_stars, compliment_hot, compliment_more, compliment_profile, compliment_cute, compliment_list, compliment_note, compliment_plain, compliment_cool, compliment_funny, compliment_writer, compliment_photos
 

### 2021-02-09 code.py
got working SVD recommender, now need to break ties with context

@@ -25,6 +25,15 @@
 
 # restaurants on yelp
 
+# json files I use:
+# user_ids
+# business_ids
+# covid
+# categories
+# cities/city_ids
+# usefulReviews.csv
+
+
 import os
 import json
 from collections import defaultdict
@@ -39,7 +48,7 @@ import numpy as np
 import math
 import pickle
 import time
-
+'''
 def collaborativeFiltering(): # code adapted from https://github.com/wwwbbb8510/baseline-rs
     ratings = pd.read_csv('processed/usefulReviews.csv', encoding='"ISO-8859-1"')
     ratings = ratings.sample(frac=0.01)
@@ -179,7 +188,7 @@ def get_top_n(predictions, n=10):
         user_ratings.sort(key=lambda x: x[1], reverse=True)
         top_n[uidLocal] = user_ratings[:n]
     return top_n
-
+'''
 def getRecommendations(uid):
     file_path = os.path.expanduser('processed/usefulReviews.csv')
     reader = Reader(line_format='user item rating timestamp', sep=',')
@@ -189,7 +198,7 @@ def getRecommendations(uid):
     trainset = data.build_full_trainset()
     algo = SVD()
     #start = time.time()
-    print('Getting recommendations...')
+    print('Getting recommendations for {} of {} restaurants in {}...'.format(name,category,city))
     algo.fit(trainset)
     #end = time.time()
     #print(end-start)
@@ -229,7 +238,6 @@ def presentRecommendations(items): # takes items from recommender and
         i += 1
     table.add_rows(rows)
     print()
-    print('Here are some {} restaurants in {} that you might like:'.format(category,city))
     print(table.draw() + '\n')
     print('Select a restaurant to see more information')
     check = False
@@ -247,7 +255,7 @@ def moreInformation(restaurant):
     if restaurant == None:
         return
     business = businesses[restaurant]
-    name = business['name']
+    restaurantName = business['name']
     address = '{}\n{}, {}'.format(business['address'],business['city'],business['state'])
     features = covidFeatures[restaurant]
     delivery = features['delivery or takeout']
@@ -261,7 +269,7 @@ def moreInformation(restaurant):
     table.set_cols_align(['l','c','c','c','c','c'])
     table.set_cols_valign(['m','m','m','m','m','m'])
     table.add_rows([['Name','Address','Stars','No. reviews','Delivery or Takeout','Grubhub'],
-    [name,address,business['stars'],business['review_count'],delivery,grubhub]])
[…]

### 2021-02-09 code.py
accommodate hours being None

@@ -220,6 +220,10 @@ def getRecommendations(uid):
 def hybrid():
     pass
     # use namsor to get info about person from name, then use to recommend 
+    # take scores produced by SVD calculations, e.g. when there's nothing to go on
+    # break ties by looking at review count, accepts credit cards, price range, attire, alcohol, 
+    # reservations, takeout, good for groups, ambience, good for kids, drive thru, wifi, parking, 
+    # caters, delivery, noise level, outdoor seating, has tv, good for meal, categories, hours
 
 def presentRecommendations(items): # takes items from recommender and 
     if items == []:
@@ -279,11 +283,12 @@ def moreInformation(restaurant):
     table.set_cols_align(['l','l'])
     rows = [['Opening Hours:','']]
     hours = business['hours']
-    for day in hours:
-        times = hours[day].replace(':0',':00')
-        rows.append([day,times])
-    table.add_rows(rows)
-    print(table.draw())
+    if hours != None:
+        for day in hours:
+            times = hours[day].replace(':0',':00')
+            rows.append([day,times])
+        table.add_rows(rows)
+        print(table.draw())
     if closed != 'FALSE':
         print()
         print('{} is closed temporarily until {}\n'.format(name,closed.split('T')[0]))

### 2021-02-10 paper/paper.tex
added modules and data files to README

@@
 When filtered for restaurants which fall into at least one of these categories, 
-20,565 had at least one check-in in 2019
+20,565 had at least one check-in in 2019, but 333 of these were temporarily closed due to covid but should be 
+open now, and 1 is still closed until June 2021. 
+19,862 restaurants had a review written about them in 2019, there were 503,979 reviews written in 2019 about these restaurants. 
 There are 2,956,316 reviews of these restaurants. 

### 2021-02-10 code.py

@@ -25,18 +25,11 @@
 
 # restaurants on yelp
 
-# json files I use:
-# user_ids
-# business_ids
-# covid
-# categories
-# cities/city_ids
-# usefulReviews.csv
 
 
-import os
+#import os
 import json
-from collections import defaultdict
+#from collections import defaultdict
 #from surprise import BaselineOnly
 from surprise import SVD
 from surprise import Dataset
@@ -44,10 +37,10 @@ from surprise import Reader
 #from surprise.model_selection import cross_validate
 import texttable
 import pandas as pd
-import numpy as np
-import math
-import pickle
-import time
+#import numpy as np
+#import math
+#import pickle
+#import time
 '''
 def collaborativeFiltering(): # code adapted from https://github.com/wwwbbb8510/baseline-rs
     ratings = pd.read_csv('processed/usefulReviews.csv', encoding='"ISO-8859-1"')
@@ -190,7 +183,7 @@ def get_top_n(predictions, n=10):
     return top_n
 '''
 def getRecommendations(uid):
-    file_path = os.path.expanduser('processed/usefulReviews.csv')
+    file_path = 'processed/usefulReviews.csv'
     reader = Reader(line_format='user item rating timestamp', sep=',')
     data = Dataset.load_from_file(file_path, reader=reader)
     #cross_validate(BaselineOnly(), data, verbose=True)
@@ -224,21 +217,30 @@ def hybrid():
     # break ties by looking at review count, accepts credit cards, price range, attire, alcohol, 
     # reservations, takeout, good for groups, ambience, good for kids, drive thru, wifi, parking, 
     # caters, delivery, noise level, outdoor seating, has tv, good for meal, categories, hours
+    # all from business.json or business_ids.json
+    # TODO
+    # implement hybrid properly, look at lectures to figure out how to do it 
+    # nearest neighbour for e.g. best night or ambience that person likes the most
+    # certain features that person seems to prefer?
+    # is SVD time-aware?
+    # comment code
+    # write paper
 
 def presentRecommendations(items): # takes items from recommender and 
     if items == []:
         print('There are no {} restaurants in {}'.format(category,city))
         return None
     table = texttable.Texttable()
-    table.set_cols_dtype(['i','t','t','t']) # specify data types
-    table.set_cols_align(['c','l','c','c']) # align columns horizontally
-    table.set_cols_valign(['m','m','m','m']) # align columns vertically
-    rows = [['Number','Name', 'City', 'Stars']] # titles of columns
[…]

### 2021-02-10 filter.py

@@ -164,9 +164,9 @@ def dropText():
     #panda = pd.read_json('processed/mostActiveReviews.json',lines=True)
     #panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
     #panda.to_csv('processed/mostActiveReviews.csv',index=False)
-    panda = pd.read_json('processed/usefulReviews.json',lines=True)
+    panda = pd.read_json('processed/recentReviews.json',lines=True)
     panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
-    panda.to_csv('processed/usefulReviews.csv',index=False)
+    panda.to_csv('processed/recentReviews.csv',index=False)
     '''
     inFile = open('processed/review.json','r')
     outFile = open('processed/review.csv','w')
@@ -229,6 +229,16 @@ def filterUsefulReviews():
     inFile.close()
     outFile.close()
 
+def filterRecentReviews():
+    inFile = open('processed/review.json','r')
+    outFile = open('processed/recentReviews.json','w')
+    for line in inFile:
+        review = json.loads(line)
+        if review['date'][:4] == '2019':
+            outFile.write(line)
+    inFile.close()
+    outFile.close()
+
 def splitReviews():
     inFile = open('processed/usefulReviews.json','r')
     closestCity = open('processed/closestCity.json','r')
@@ -275,6 +285,20 @@ def formatCovidFeatures():
     inFile.close()
     outFile.close()
 
+def closedCovid():
+    lastFile = open('processed/covid.json','r')
+    covid = json.load(lastFile)
+    lastFile.close()
+    inFile = open('processed/business.json','r')
+    outFile = open('processed/open2019Business.json','w')
+    for line in inFile:
+        business = json.loads(line)
+        ID = business['business_id']
+        if covid[ID]['Temporary Closed Until'] == 'FALSE':
+            outFile.write(line)
+    inFile.close()
+    outFile.close()
+
 
 
 #ids = getIDs()
@@ -289,15 +313,28 @@ def formatCovidFeatures():
 #mostReviews()
 #filterMostActiveReviews()
 #uniqueRestaurants()
-#dropText()
+dropText()
 #splitCities()
 #splitIDs()
 #filterUsefulReviews()
+#filterRecentReviews()
 #splitReviews()
 #getCategories()
-formatCovidFeatures()
+#formatCovidFeatures()
+#closedCovid()
 
[…]

### 2021-02-10 paper/paper.tex
added processed data I use

@@
 \subsection{Input interface}
+The system has a command line based interface, so all user input of the input is given via the command line. 
+The system recognises the active user by their unique user ID, some example IDs are provided in the README.txt 
+file for testing and demonstration purposes. 
+Only explicit user data is gathered in order to make the system more explainable and transparent.
+
 The input interface is the command line.

### 2021-02-10 code.py

@@ -244,7 +244,13 @@ def presentRecommendations(items): # takes items from recommender and
         i += 1
     table.add_rows(rows)
     print()
-    print(table.draw() + '\n')
+    print(table.draw())
+    print('''
+These recommendations are based on: 
+    1. The opinions of other users with similar preferences to you
+    2. The city you said you were closest to ({})
+    3. The category of food you selected ({})
+    '''.format(city,category))
     print('Select a restaurant to see more information')
     check = False
     while not check:

### 2021-02-10 paper/paper.tex
added ability to give rating

--- 2021-02-10 paper/paper.tex
@@
 
-\author{\IEEEauthorblockN{Thomas Butterfield}
+\author{\IEEEauthorblockN{lrfk99}
 \IEEEauthorblockA{\textit{dept. Computer Science} \\
 \textit{Durham University}\\
-Durham, United Kingdom \\
-thomas.butterfield@durham.ac.uk}
+Durham, United Kingdom}
 }
@@
 \subsection{Input interface}
-The input interface is the command line.
+The system has a command line based interface, so all user input of the input is given via the command line. 
+The system recognises the active user by their unique user ID, some example IDs are provided in the 'README.txt' 
+file for testing and demonstration purposes. 
+Only explicit user data is gathered in order to make the system more explainable and transparent, 
+and users are made aware of which data is collected, how it is collected, and for which purposes. 
+This data includes the user's nearest city (of the 10 available) and the type of food that 
+they are in the mood for. 
+The user is able to change these choices from the main menu, as well as switch to a different user 
+for testing purposes. 
+When the user is presented with a list of recommendations, they can then choose one that they want 
+to see more information about and from here they can leave a rating, return to the list, or return to 
+the main menu. 
+User profiles can be updated by giving a rating of a restaurant through the system. 
+Any rating given will affect future recommendations given by the system to that user 
+and to other similar users.
+
 The program offers users opportunities to input information about themselves,

### 2021-02-10 code.py

@@ -184,7 +184,7 @@ def get_top_n(predictions, n=10):
 '''
 def getRecommendations(uid):
     file_path = 'processed/usefulReviews.csv'
-    reader = Reader(line_format='user item rating timestamp', sep=',')
+    reader = Reader(line_format='user item rating', sep=',')
     data = Dataset.load_from_file(file_path, reader=reader)
     #cross_validate(BaselineOnly(), data, verbose=True)
 
@@ -230,6 +230,13 @@ def presentRecommendations(items): # takes items from recommender and
     if items == []:
         print('There are no {} restaurants in {}'.format(category,city))
         return None
+    print('''
+These recommendations are based on: 
+    1. Your previous ratings of these restaurants (if any)
+    2. The opinions of other users with similar preferences to you
+    3. The city you said you were closest to ({})
+    4. The category of food you selected ({})
+    '''.format(city,category))
     table = texttable.Texttable()
     table.set_cols_dtype(['i','t','t','t','t']) # specify data types
     table.set_cols_align(['c','l','c','c','c']) # align columns horizontally
@@ -243,15 +250,9 @@ def presentRecommendations(items): # takes items from recommender and
         rows.append([i,business['name'],business['city'],business['stars'],predRating])
         i += 1
     table.add_rows(rows)
-    print()
     print(table.draw())
-    print('''
-These recommendations are based on: 
-    1. The opinions of other users with similar preferences to you
-    2. The city you said you were closest to ({})
-    3. The category of food you selected ({})
-    '''.format(city,category))
-    print('Select a restaurant to see more information')
+    print()
+    print('Select a restaurant to see more information or rate it: (1-{})'.format(len(items)))
     check = False
     while not check:
         choice = input('Your Choice > ').strip()
@@ -308,40 +309,58 @@ def moreInformation(restaurant,predRating):
         print()
     print('''
 What do you want to do?
-1. See the list again
-2. Return to main menu
-    ''')
+1. Give a rating of {}
+2. See the list again
+3. Return to main menu
+    '''.format(restaurantName))
     while True:
         choice = input('Your Choice > ').strip()
         try:
             choice = int(choice)
             if choice == 1:
+                print('''
+How many stars do you want to give {}? (1-5)
+                '''.format(restaurantName))
+                while True:
+                    stars = input('Your Rating > ').strip()
+                    try:
+                        stars = int(stars)
+                        if 1 <= stars <= 5:
+                            reviews = open('processed/usefulReviews.csv','a')
[…]

### 2021-02-10 filter.py

@@ -161,12 +161,9 @@ def uniqueRestaurants():
         print(len(businesses),businesses[max(businesses,key=businesses.get)])
 
 def dropText():
-    #panda = pd.read_json('processed/mostActiveReviews.json',lines=True)
-    #panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
-    #panda.to_csv('processed/mostActiveReviews.csv',index=False)
-    panda = pd.read_json('processed/recentReviews.json',lines=True)
-    panda = panda.drop(['review_id','useful','funny','cool','text'],axis=1)
-    panda.to_csv('processed/recentReviews.csv',index=False)
+    panda = pd.read_json('processed/usefulReviews.json',lines=True)
+    panda = panda.drop(['review_id','useful','funny','cool','text','date'],axis=1)
+    panda.to_csv('processed/usefulReviews.csv',index=False)
     '''
     inFile = open('processed/review.json','r')
     outFile = open('processed/review.csv','w')
@@ -327,14 +324,3 @@ dropText()
 
 # filter by time
 # get number of reviews in last year?
-
-'''
-inFile = open('processed/recentReviews.json','r')
-business = {}
-for line in inFile:
-    ID = json.loads(line)['business_id']
-    if ID not in business:
-        business[ID] = None
-print(len(business))
-inFile.close()
-'''

### 2021-02-10 code.py
deleted unused code and imports

@@ -27,9 +27,7 @@
 
 
 
-#import os
 import json
-#from collections import defaultdict
 #from surprise import BaselineOnly
 from surprise import SVD
 from surprise import Dataset
@@ -37,151 +35,7 @@ from surprise import Reader
 #from surprise.model_selection import cross_validate
 import texttable
 import pandas as pd
-#import numpy as np
-#import math
-#import pickle
-#import time
-'''
-def collaborativeFiltering(): # code adapted from https://github.com/wwwbbb8510/baseline-rs
-    ratings = pd.read_csv('processed/usefulReviews.csv', encoding='"ISO-8859-1"')
-    ratings = ratings.sample(frac=0.01)
-    ratings_training = ratings.sample(frac=0.7)
-    ratings_test = ratings.drop(ratings_training.index)
-    rating_mean = ratings_training.groupby(['business_id'], as_index=False, sort=False).mean().rename(columns={'rating':'rating_mean'})[['business_id','rating_mean']]
-    adjusted_ratings = pd.merge(ratings_training,rating_mean,on='business_id',how='left',sort=False)
-    adjusted_ratings['rating_adjusted']=adjusted_ratings['rating']-adjusted_ratings['rating_mean']
-    adjusted_ratings.loc[adjusted_ratings['rating_adjusted']==0,'rating_adjusted'] = 1e-8
-    start = time.time()
-    w_matrix = build_w_matrix(adjusted_ratings,True)
-    end = time.time()
-    print(end-start)
-    print('built w_matrix')
-    #predict(uid,iid,w_matrix,adjusted_ratings,rating_mean)
-    recommended_restaurants = recommend(uid,w_matrix,adjusted_ratings,rating_mean)
-    presentRecommendations(recommended_restaurants)
-    #items = []
-    #for i in range(8):
-    #    items.append(recommended_restaurants['business_id'].iloc[i])
-    #print(items)
-    #presentRecommendations(items)
-    
-    print(recommended_restaurants)
-    end = time.time()
-    print(end-start)
 
-def build_w_matrix(adjusted_ratings, load_existing_w_matrix):
-    w_matrix_columns = ['business_1', 'business_2', 'weight']
-    w_matrix = pd.DataFrame(columns=w_matrix_columns)
-    if load_existing_w_matrix:
-        with open('processed/w_matrix.pkl','rb') as input:
-            w_matrix = pickle.load(input)
-        input.close()
-    else:
[…]

### 2021-02-11 paper/paper.tex
fleshed out introduction section of paper

@@
 \subsection{Domain of application}
-The domain of this application is a recommender system for bars, using the Yelp dataset.
-Decide whether I want to define the time-span for the date, 
-depends on if there is too much data from all available time period.
-I am also taking into accounts Covid-19 data when generating recommendations.
+The domain of this hybrid recommender system is restaurants, with a focus on trying new restaurants. 
+This system is built around restaurants in a particular metropolitan area which serve a particular 
+type of food.
 
 \subsection{Related work review}
-Some related work
+Hybrid recommender systems which combine Knowledge-based (KB) and Collaborative-filtering (CF) techniques 
+are designed to combine the ability of KB techniques to give generic recommendations to new users with 
+small profiles and the strength of CF to find peer users with unexpected shared preferences. 
+This combination of techniques is particularly well suited to restaurants since a user may often be in the 
+mood for a specific type of food, such as Italian or Chinese, so they only want to see suggestions of that 
+type or restaurant, which is where the Knowledge-based recommender is helpful. 
+A Collaborative-filtering algorithm will then be able to order all the restaurants which match this 
+explicit criteria by how the user should rate them, based on the opinions of other similar users, 
+helping them to discover new restaurants to try. 
+These techniques are used in the Entree restaurant recommender system \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
+In particular, Burke's 1999 paper \cite{burke1999integrating} is the first to discuss specifically using these 
+two techniques in the same cascaded style that I implement here.
 
 \subsection{Purpose/Aim}
-The purpose of this application is to give suitable suggestions for a user to go to.
-Using information about the user such as their location and personal preferences.
+The purpose of this application is to give suitable suggestions of restaurants for a user to go to. 
+The recommended restaurants should be in the metropolitan area which the user specifies, and serve the 
+type of food that the user asks for. 
+As well as fitting this explicit criteria, the suggested restaurants should be similar to other restaurants 
+that the user has rated highly in the past and which other users with similar preferences also like. 
+The system will allow the user to add ratings of the restaurants they are recommended, and these ratings will 
+affect future suggestions the system gives for that user and other similar users.
 
@@
 \subsection{Data description}
+includes 10 metropolitan areas in the USA and Canada (20,565 restaurants) as well as given consideration 
+to Covid-19 data when  which is included in the dataset. 
 The data I am using is taken from the Yelp dataset,
@@
 
-The program offers users opportunities to input information about themselves,
-as well as make choices from a selection of items which the system provides or suggests.
-
 \subsection{Recommendation algorithm}
-What algorithm
+I used a model-based approach to Collaborative Filtering based on matrix factorisation. 
+The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
+In my program, I fit the algorithm on a training set, and then for every restaurant of the specified category 
+and in the specified city, I predict what the user would rate that restaurant. 
+Then I sort those restaurants by their score and take the top 8 to present to the user.
 
 \subsection{Output interface}
-The output interface is the command line.
-The system can output recommendations for bars which the user might like,
-as well as information about how the system works and why certain suggestions were made,
-at the user's request.
+Recommendations and prediction scores are presented to the user via the command line interface and using an ASCII table. 
+I used a python module called texttable for this, since it makes the output very clear and visually appealing. 
+I chose to output the top 8 recommendations since I think this is enough variety and choice for the user to find 
+a restaurant to try, without there being too many options which might overwhelm the user. 
+If I presented more than 8 recommendations at a time, the table wouldn't fit in a single terminal output window and 
+would require the user to scroll around to read different parts of the output. 
+Since the user will be interacting with the system on a laptop or desktop computer through a command line interface, 
+an ASCII table is the most visually attractive way to present information, and the white text on a (typically) black 
+background will help accommodate users who are colourblind or who require a high-contrast interface. 
+The interface is clear enough to understand that users of all ages are able to use the system, and if a user 
+requires the text to be larger then they can increase the text size of their terminal. 
+Along with the recommended restaurants, an explanation of how these results are calculated is also provided, as 
+well as the prediction values themselves so the user can understand why they are being shown these suggestions. 
+This, along with an option in the main menu to view information about the data the system uses and how it is used, 
+further adds to the explainability and transparency.
 

### 2021-02-11 code.py

@@ -211,7 +211,7 @@ How is it collected?
 
 For what purpose?
 
-    In order for us to suggest restaurants that you might like.
+    In order for us to suggest more relevant restaurants that you might like.
 
     ''')
     input('Press enter to return to main menu ')

### 2021-02-11 paper/paper.tex
wrote data description section

@@
 \documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
 \usepackage{cite}
@@
 \subsection{Data description}
-includes 10 metropolitan areas in the USA and Canada (20,565 restaurants) as well as given consideration 
-to Covid-19 data when  which is included in the dataset. 
-The data I am using is taken from the Yelp dataset,
-it includes user reviews of different businesses and services in a specific location.
-There are 10 cities included in the dataset: Montreal, Calgary, Toronto, Pittsburgh, Charlotte, 
-Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland. 
-These user reviews include ratings, text feedback, and other such information.
+The data for my recommender system was generated from the Yelp Open Dataset \cite{YelpDataset}. 
+This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
+The dataset includes approximately 8,000,000 reviews of 200,000 businesses by 2,000,000 users, as well as 
+additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
+The dataset also includes approximately 200,000 photos but I these are not relevant to my application, since 
+the interface is command line only, so they were discarded. 
+Every businesses in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
+(Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.
 
 \subsection{Data preparation and feature selection}
+The business data was filtered down to businesses that contain 'Restaurants' as one of their listed 
+categories
+The ratings data was 
+(20,565 restaurants)
 Prepare data
@@
 by city. 
+Mention Netflix Prize and proposition of SVD, use lecture slides explanation and surprise documentation.
 

### 2021-02-11 paper/paper.tex
renamed files and cleaned up comments

@@
 additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
-The dataset also includes approximately 200,000 photos but I these are not relevant to my application, since 
+There are 3 sections of the dataset, with 5 JSON files in the main dataset, 1 in the photos section, and 
+1 in the covid section. 
+The $business.json$ file contains business data including location data, attributes, and categories. 
+$review.json$ contains full review text data including the $user\_id$ that wrote the review and the $business\_id$ 
+that the review is written for. 
+The file $user.json$ holds every user's first name along with other data adn metadata which I don't use \cite{YelpDataset}. 
+The data files $checkin.json$ and $tip.json$ are not used in my system. 
+The dataset also includes approximately 200,000 photos but these are not relevant to my application, since 
 the interface is command line only, so they were discarded. 
-Every businesses in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
+The covid related data from $covid\_features.json$ contains additional information about each business with regards to 
+the measures they have taken due to the pandemic, such as offering delivery or takeout options or a 
+special message for customers. 
+Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
 (Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.
@@
 \subsection{Data preparation and feature selection}
+The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
+which does not lend well to fast lookups by the main program. 
+Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
+be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
+$user\_id$ depending on the case. 
+This allows for extremely fast data lookups in the system. 
+It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
+I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
+categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
+Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
+Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
+never be recommended a restaurant which doesn't fit at least one of these categories. 
+I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
+data was available). 
+This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
+metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
+The list of reviews pertaining to these restaurants was then filtered down
+I also split by nearest city, and filtered only useful reviews
+This was done using the $categories$ attribute of the $business$ JSON file.
+
 The business data was filtered down to businesses that contain 'Restaurants' as one of their listed 
 categories
+getIDs
+filterReviews
+lastCheckin
+recentCheckin
+toCSV
+CovidDuplicates
+getUserIDs
+mostActiveUsers
+numberRestaurantReviews
+mostReviews
+filterMostActiveReviews
+uniqueRestaurants
+dropText
+splitCities
+splitIDs
+filterUsefulReviews
+filterRecentReviews
+splitReviews
+getCategories
+formatCovidFeatures
+closedCovid
+
 The ratings data was 
@@
 \subsection{Recommendation techniques/algorithms}
-The first recommender system is collaborative filtering.
+The first recommender system is collaborative filtering. reference https://surpriselib.com/
 I am using item-based collaborative filtering, since the number of users is larger than the number of items 
@@
 \subsection{Recommendation algorithm}
+The idea is to find restaurants in a new city which are similar to those in another city that the user likes, just like Entree. 
 I used a model-based approach to Collaborative Filtering based on matrix factorisation. 
@@
 What could I do in the future \cite{5284958}. 
+Implement the idea of particular categories being closer some than others. 
+Specify price range and/or consider price when comparing similarities of restaurants.
 

### 2021-02-11 RS.py

@@ -1,16 +1,5 @@
 
-# first step, look at dataset, decide what services to include, start preprocessing data
-# then build architecture/scheme
-
-# personalised recomendations, not just demographic or generic
-# look at the individual user and their preferences and needs, context, interests
-# justify choice of models and justify all choices
 # you must compare the other systems using evaluation metrics but actual performance of recommender system is not important
-
-# command line interface
-# allow user to specify their name (or other identifying property), so model allows distinct users and gives personalised recomendations
-
-# marked on explainability
 
 # video is about how well you can present your product. Voice over video
 # explain what it is, how it works, display the best of its features, demo main functionality
@@ -23,9 +12,7 @@
 #   limitations
 # in order to get an idea of common or good or useful combinations of systems
 
-# restaurants on yelp
-
-
+# restaurants
 
 import json
 #from surprise import BaselineOnly
@@ -37,7 +24,7 @@ import texttable
 import pandas as pd
 
 def getRecommendations(uid):
-    file_path = 'processed/usefulReviews.csv'
+    file_path = 'processed/reviews.csv'
     reader = Reader(line_format='user item rating', sep=',')
     data = Dataset.load_from_file(file_path, reader=reader)
     #cross_validate(BaselineOnly(), data, verbose=True)
@@ -51,7 +38,8 @@ def getRecommendations(uid):
     #print(end-start)
     suggestions = {}
     for iid in iids:
-        if iid in categories and category in categories[iid]: # include all reviews?
+        if iid in businesses and category in businesses[iid]['categories']:
+            #if iid in categories and category in categories[iid]: # include all reviews?
             pred = algo.predict(uid,iid)#,verbose=True)#,r_ui=4, verbose=True)
             suggestions[pred[1]] = pred[3]
     suggestions = sorted(suggestions.items(), key=lambda item: item[1],reverse=True)
@@ -180,7 +168,7 @@ How many stars do you want to give {}? (1-5)
                     try:
                         stars = int(stars)
                         if 1 <= stars <= 5:
-                            reviews = open('processed/usefulReviews.csv','a')
+                            reviews = open('processed/reviews.csv','a')
                             reviews.write('{},{},{}\n'.format(uid,restaurant,stars))
                             reviews.close()
                             break
@@ -245,8 +233,7 @@ q. Quit
[…]

### 2021-02-11 filter.py

@@ -296,8 +296,6 @@ def closedCovid():
     inFile.close()
     outFile.close()
 
-
-
 #ids = getIDs()
 #filterRelevant()
 #lastCheckin()
@@ -310,7 +308,7 @@ def closedCovid():
 #mostReviews()
 #filterMostActiveReviews()
 #uniqueRestaurants()
-dropText()
+#dropText()
 #splitCities()
 #splitIDs()
 #filterUsefulReviews()
@@ -319,8 +317,3 @@ dropText()
 #getCategories()
 #formatCovidFeatures()
 #closedCovid()
-
-# user features: user_id, name, review_count, yelping_since, useful, funny, cool, elite, friends, fans, average_stars, compliment_hot, compliment_more, compliment_profile, compliment_cute, compliment_list, compliment_note, compliment_plain, compliment_cool, compliment_funny, compliment_writer, compliment_photos
-
-# filter by time
-# get number of reviews in last year?

### 2021-02-11 RS.py
renamed city json files

@@ -143,10 +143,10 @@ def moreInformation(restaurant,predRating):
         print(table.draw())
     if closed != 'FALSE':
         print()
-        print('{} is closed temporarily until {}\n'.format(name,closed.split('T')[0]))
-    if message != 'FALSE':
+        print('{} is/was closed temporarily until {}\n'.format(restaurantName,closed.split('T')[0]))
+    if message != 'FALSE' and message != '':
         print()
-        print('Covid message from {}:'.format(name))
+        print('Covid message from {}:\n'.format(restaurantName))
         print(message)
         print()
     print('''
@@ -306,7 +306,7 @@ if __name__ == '__main__':
     uid, name = getID()
     category = getCategory()
     city = whichCity()
-    cityRestaurants = open('processed/cities/'+city+'_ids.json','r')
+    cityRestaurants = open('processed/cities/'+city+'.json','r')
     iids = json.load(cityRestaurants)
     cityRestaurants.close()
     items = getRecommendations(uid)
@@ -323,7 +323,7 @@ if __name__ == '__main__':
             category = getCategory()
         elif choice == 3:
             city = whichCity()
-            cityRestaurants = open('processed/cities/'+city+'_ids.json','r')
+            cityRestaurants = open('processed/cities/'+city+'.json','r')
             iids = json.load(cityRestaurants)
             cityRestaurants.close()
         elif choice == 4:

### 2021-02-11 paper/paper.tex
changed demo user IDs to better ones

@@
 never be recommended a restaurant which doesn't fit at least one of these categories. 
+I chose to use the top 12 categories because it's enough to provide a wide variety of options to the user without 
+being so many that the user is overwhelmed by the amount. 
 I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
@@
 metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
-The list of reviews pertaining to these restaurants was then filtered down
-I also split by nearest city, and filtered only useful reviews
-This was done using the $categories$ attribute of the $business$ JSON file.
+All other datasets were filtered to only data pertaining to these restaurants, and the reviews were further filtered 
+to only those which had 5 or more 'useful' votes by other users. 
+This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
+recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
+This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) and structured as follows: 
+($User\_ID$, $Business\_ID$, $Rating$).
 
-The business data was filtered down to businesses that contain 'Restaurants' as one of their listed 
-categories
-getIDs
-filterReviews
-lastCheckin
-recentCheckin
-toCSV
-CovidDuplicates
-getUserIDs
-mostActiveUsers
-numberRestaurantReviews
-mostReviews
-filterMostActiveReviews
-uniqueRestaurants
-dropText
-splitCities
-splitIDs
-filterUsefulReviews
-filterRecentReviews
-splitReviews
-getCategories
-formatCovidFeatures
-closedCovid
-
-The ratings data was 
-(20,565 restaurants)
-Prepare data
-Of the 209,393 businesses in the dataset, 168,903 are still open. 
-There are 43,965 restaurants which are open and contain 'Restaurants' as one of the categories. 
-I then picked the top 12 most popular categories of restaurant: American, Mexican, Italian, Chinese, 
-Seafood, Japanese, Canadian, Mediterranean, Indian, Thai, Middle Eastern, Vietnamese (in that order). 
-When filtered for restaurants which fall into at least one of these categories, 
 20,565 had at least one check-in in 2019, but 333 of these were temporarily closed due to covid but should be 

### 2021-02-11 filter.py

@@ -296,6 +296,20 @@ def closedCovid():
     inFile.close()
     outFile.close()
 
+def getDemoIDs():
+    inFile = open('processed/usefulReviews.json','r')
+    uids = {}
+    for line in inFile:
+        review = json.loads(line)
+        uid = review['user_id']
+        if uid in uids:
+            uids[uid] += 1
+        else:
+            uids[uid] = 1
+    inFile.close()
+    print(sorted(uids.items(), key=lambda item: item[1]))
+    print(len(uids))
+
 #ids = getIDs()
 #filterRelevant()
 #lastCheckin()
@@ -317,3 +331,4 @@ def closedCovid():
 #getCategories()
 #formatCovidFeatures()
 #closedCovid()
+#getDemoIDs()

### 2021-02-11 paper/paper.tex
added lots of equations to algorithm section

@@
 \subsection{Related work review}
-Hybrid recommender systems which combine Knowledge-based (KB) and Collaborative-filtering (CF) techniques 
+Hybrid recommender systems which combine knowledge-based (KB) and collaborative-filtering (CF) techniques 
 are designed to combine the ability of KB techniques to give generic recommendations to new users with 
@@
 mood for a specific type of food, such as Italian or Chinese, so they only want to see suggestions of that 
-type or restaurant, which is where the Knowledge-based recommender is helpful. 
-A Collaborative-filtering algorithm will then be able to order all the restaurants which match this 
+type or restaurant, which is where the knowledge-based recommender is helpful. 
+A collaborative-filtering algorithm will then be able to order all the restaurants which match this 
 explicit criteria by how the user should rate them, based on the opinions of other similar users, 
@@
 \subsection{Data preparation and feature selection}
+The entire Yelp dataset contains a lot of data which is not relevant to my domain of restaurants, so much data 
+preparation is required to save computation time and in order to provide more accurate predictions. 
 The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
@@
 metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
-All other datasets were filtered to only data pertaining to these restaurants, and the reviews were further filtered 
-to only those which had 5 or more 'useful' votes by other users. 
+All other datasets were filtered to only data pertaining to these restaurants, with duplicates in the 
+covid dataset removed. 
+The reviews were further filtered to only those which had 5 or more 'useful' votes by other users. 
 This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
 recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
-This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) and structured as follows: 
-($User\_ID$, $Business\_ID$, $Rating$).
-
-20,565 had at least one check-in in 2019, but 333 of these were temporarily closed due to covid but should be 
-open now, and 1 is still closed until June 2021. 
-19,862 restaurants had a review written about them in 2019, there were 503,979 reviews written in 2019 about these restaurants. 
-There are 2,956,316 reviews of these restaurants. 
-481,073 tips for these restaurants. 
-1,968,703 users. 
-None of these restaurants are temporarily closed according to the covid features data. 
-20,573 restaurants listed with covid features, 4 restaurants which make up 8 duplicate entries 
-(4 have 3 entries), the only difference between their entries being in the 
-"Virtual Services Offered" section. 
-Allow to pick user ID of 10 users with most reviews made, to avoid cold start problem
-4 users have over 10,000 reviews, the most is 14,455 (Victor, 8k3aO-mPeyhbR5HUucA5aA). 
-I actually want to select the most reviews of the places I am accepting, restaurants. 
-When taking the most reviews of restaurants in teh dataset, a threshold of 600 reviews gives a list of 10 users, 
-with the most being 1747. 
-I then checked each of these for how many different restaurants they had reviewed and the greatest number of times 
-they had reviewed a single restaurant to check that none of them were just users who reviewed a small number of 
-restaurants a huge amount of times each. 
-All 10 users passed this criteria, with even the least active user having reviewed 496 unique restaurants, which will 
-certainly avoid any cold-start problems. 
-I split the businesses into 10 groups, by the city which they are closest to, I did this by using latitude and 
-longitude rather than State or City since this does not give actual distance and restaurants close to a border 
-could then be mis-categorised. 
-Since there are 4,223,201 reviews of restaurants I filtered them by removing any which had zero 'useful' votes. 
-When filtered to reviews with at least 1 'useful' vote, there are 1,735,732 and when at least 2 votes, 836,929
-At least 3 votes, 462,412, at least 5 votes 190,902. 
-51,737 with at least 10 votes
-
-Top categories: American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, Mediterranean, Indian, Thai, Middle Eastern, Vietnamese
-
-The entire Yelp dataset is huge and much of it is not necessary for my domain of bars.
-As such I prepared the data by eliminating any data not relevant to my domain.
-I selected features such as user ratings and a particular user's average rating,
-since if a user typically rates places they go highly, but rates a particular bar low,
-this is more significant than a user who always rates places low.
+This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) with unnecessary data 
+removed and structured as follows: ($User\_ID$, $Business\_ID$, $Rating$).
 
 \subsection{Hybrid scheme}
-Which two algorithms
-A hybrid scheme is a good way to design a recommender system, since you can get the
-best of both algorithms if done properly. 
-I am using a Cascade Hybrid Recommender System, since this has been shown to accurate predictions 
-in my domain of restaurant recommendations \cite{burke2007hybrid}. 
-The two systems are Collaborative Filtering and Content Based.
-
-Meaning better recommendations than either algorithm could achieve individually.
-Cascade with knowledge-based and collaborative
+The two personalised recommender systems I implement are knowledge-based and collaborative-filtering, and 
+the hybrid scheme is cascade. 
+A cascade hybrid scheme works particularly well in my domain of restaurants with an emphasis on providing accurate 
+predictions of restaurants which a user may like in a city they have never been to before, based on the restaurants 
+they like in their home-city. 
+This is because the knowledge-based system can filter the restaurants by location and by the type of food they serve, 
+while the collaborative-filtering system is able to find restaurants which are similar to the ones the user has 
+previously rated highly. 
+So when these two systems are combined in a cascade style with the knowledge-based system as the primary recommender 
+and the collaborative-filtering system as the secondary the hybrid system can find hidden similarities between 
+restaurants in different cities very accurately. 
+This method of hybrid recommender system (HRS) has been implemented as a restaurant recommender in the Entree 
+system, giving very accurate results \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}.
 
 \subsection{Recommendation techniques/algorithms}
-The first recommender system is collaborative filtering. reference https://surpriselib.com/
-I am using item-based collaborative filtering, since the number of users is larger than the number of items 
-(1,968,703 users vs 209,393 businesses in the entire dataset), this gives greater accuracy of predictions. 
-In this case the list of available items does not change, so a user-based method would be unnecessary. 
-One of the requirements for this system is justifiability, and item-based methods make it much easier to 
-explain why certain predictions were made. 
-Item-item collaborative filtering looks for restaurants that are similar, in terms of how people rate them, 
-to the restaurants that the user has already rated and recommend the most similar restaurants. 
-Filtering reviews by city and then using CF won't work because the active user may not have ever reviewed a restaurant 
-in that city, so there is no way to assess similar restaurants to ones they have already rated. 
-Instead I carried out CF on all high quality reviews and produce an ordering of recommendations, which I then filter 
-by city. 
-Mention Netflix Prize and proposition of SVD, use lecture slides explanation and surprise documentation.
-
-I am using a weighted mixed combination of these two systems in order to produce the results
-of my hybrid recommender system.
+The first prediction method used in this hybrid system is a knowledge-based (KB) approach, which relies on knowledge 
+about each of the restaurants based on its location and the type of food that it serves. 
+This approach completely eliminates the cold-start problem, a new user will still receive relevant recommendations 
+for the city they are in and the type of food they in the mood for when using a KB system. 
+The second prediction method is a collaborative-filtering (CF) approach, specifically Single Value Decomposition (SVD) 
+matrix factorisation. 
+This approach was chosen because it scales well with large datasets, which is important since my dataset has 
+~120,000 entries. 
+The CF technique here also introduces an element of variability in the recommendations, such that the user 
+may not receive the same ordering of suggestions when they run the system multiple times with the same 
+KB specifications, helping the user to try new things. 
+This system allows the user to make ratings through the user interface, providing the ability to experiment with 
+the system in terms of how new ratings affect the current user's recommendations as well as those of other similar users.
 
@@
 \subsection{Recommendation algorithm}
+The collaborative-filtering aspect of the algorithm is carried out by first converting the ratings data to a 
+matrix with $User\_ID$ as the rows and $Business\_ID$ as the columns. 
+If a user has not rated a restaurant then the rating in that cell is set to 0. 
+The prediction $\widehat{r}_{ui}$ (user $u$'s rating of item $i$) is set as:
+\begin{equation}
+    \widehat{r}_{ui} = \mu + b_u + b_i + q_i^Tp_u
+\end{equation}
+Where the parameters are learnt by minimising the following regularised squared error: 
+\begin{equation}
+    \sum_{r_{ui}\in{R_{train}}} (r_{ui} - \widehat{r}_{ui})^2 + \lambda(b_i^2 + b_u^2 + ||q_i||^2 + ||p_u||^2)
+\end{equation}
+This minimisation is performed by stochastic gradient descent:
+\begin{align}
+    \begin{split}
+    & b_u \gets b_u + \gamma(e_{ui} - \lambda b_u) \\
+    & b_i \gets b_i + \gamma(e_{ui} - \lambda b_i) \\ 
+    & p_u \gets p_u + \gamma(e_{ui} \cdot q_i - \lambda p_u) \\ 
+    & q_i \gets q_i + \gamma(e_{ui} \cdot p_u - \lambda q_i) \\
+    \end{split}
+\end{align}
+where $e_{ui} = r_{ui} - \widehat{r}_{ui}$. 
+These steps are performed over all the ratings of the trainset and repeated 20 times \cite{SVD}.
+
 The idea is to find restaurants in a new city which are similar to those in another city that the user likes, just like Entree. 
-I used a model-based approach to Collaborative Filtering based on matrix factorisation. 
+I used a model-based approach to collaborative-filtering based on matrix factorisation. 
 The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
@@
 further adds to the explainability and transparency. 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.5\linewidth]{output.png}
+    \caption{System output}
+    \label{fig:output}
+\end{figure}
 
@@
 Specify price range and/or consider price when comparing similarities of restaurants. 
+Include demographic information
 

### 2021-02-11 RS.py

@@ -79,7 +79,7 @@ These recommendations are based on:
     3. The city you said you were closest to ({})
     4. The category of food you selected ({})
     '''.format(city,category))
-    table = texttable.Texttable()
+    table = texttable.Texttable(max_width=100)
     table.set_cols_dtype(['i','t','t','t','t']) # specify data types
     table.set_cols_align(['c','l','c','c','c']) # align columns horizontally
     table.set_cols_valign(['m','m','m','m','m']) # align columns vertically
@@ -120,7 +120,7 @@ def moreInformation(restaurant,predRating):
     grubhub = 'Yes' if grubhub=='TRUE' else 'No'
     message = features['Covid Banner']
     closed = features['Temporary Closed Until']
-    table = texttable.Texttable()
+    table = texttable.Texttable(max_width=100)
     table.set_cols_dtype(['t','t','t','t','t','t','t'])
     table.set_cols_align(['l','c','c','c','c','c','c'])
     table.set_cols_valign(['m','m','m','m','m','m','m'])

### 2021-02-11 paper/paper.tex
rewriting evaluation methods

@@
 1 in the covid section. 
-The $business.json$ file contains business data including location data, attributes, and categories. 
+The $business.json$ file contains business data including location data, attributes, and cuisine. 
 $review.json$ contains full review text data including the $user\_id$ that wrote the review and the $business\_id$ 
@@
 I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
-categories in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
+cuisines in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
 Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
-Since the user selects one of these 12 options for categories of restaurants they want to find, they will 
+Since the user selects one of these 12 options for cuisines they want to find, they will 
 never be recommended a restaurant which doesn't fit at least one of these categories. 
-I chose to use the top 12 categories because it's enough to provide a wide variety of options to the user without 
+I chose to use the top 12 cuisines because it's enough to provide a wide variety of options to the user without 
 being so many that the user is overwhelmed by the amount. 
@@
 \subsection{Evaluation methods}
-How to evaluate
+The evaluation of the system was done using offline experiments. 
+The accuracy of ratings predictions was evaluated using Mean Absolute Error (MAE), which is calculated by the 
+following equation:
+\begin{equation}
+    MAE = \frac{1}{|\mathcal{T}|} \sum_{(u,i)\in{\mathcal{T}}} {|\widehat{r}_{ui} - r_{ui}|}
+\end{equation}
+This metric is appropriate for this RS since it can handle large errors on a few items, which may occur in this system 
+due to some rare skewed ratings. 
+The MAE of this system is around 0.99. 
+Measuring the accuracy of usage predictions is done using precision, calculated by: 
+\begin{equation}
+    \frac{TP}{TP+FP}
+\end{equation}
+where TP and FP correspond to True Positive and False Positive respectively. 
+This is because the number of recommendations is predefined and not very large (8 items). 
+The precision of this system is 0.125, since 1 out of the 8 items is a true positive. 
+Novelty, diversity, coverage
 
@@
 These steps are performed over all the ratings of the trainset and repeated 20 times \cite{SVD}. 
-
-The idea is to find restaurants in a new city which are similar to those in another city that the user likes, just like Entree. 
-I used a model-based approach to collaborative-filtering based on matrix factorisation. 
-The specific model is Single Value Decomposition (SVD), which was popularised by Simon Funk during the Netflix Prize. 
-In my program, I fit the algorithm on a training set, and then for every restaurant of the specified category 
-and in the specified city, I predict what the user would rate that restaurant. 
-Then I sort those restaurants by their score and take the top 8 to present to the user.
+The knowledge-based aspect of the algorithm produces recommendations based on the explicit location and cuisine 
+preferences of the user. 
+These are then cascaded down to the CF algorithm, which assigns each a predicted rating based on the ratings data. 
+The 8 recommendations with the highest predicted rating are then presented to the user.
 
@@
 \subsection{Comparison against baseline implementation}
-Compare vs generic suggestions
+Cross validation with 5 folds was carried out on both the SVD system and a baseline implementation. 
+The 5 Mean Absolute Errors for the SVD system had a mean of 0.9962 and a standard deviation of 0.0033. 
+The MAE results for the baseline implementation had a mean of 1.0091 and a standard deviation of 0.0056. 
+This shows that when compared to a baseline implementation, the SVD system does perform better, generating more 
+accurate predictions, however the difference is not particularly large.
 
 \subsection{Comparison against hybrid recommenders in related studies}
-Read some papers
+The Entree restaurant recommender system is an example of a cascade style hybrid recommender system which 
+uses both knowledge-based and collaborative-filtering methods to help users find a restaurant they might like in 
+a city they've never been to before \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
+This is very similar to the way my system works, with the biggest difference being the vastly greater amount of 
+knowledge engineering that has gone into Entree. 
+This additional knowledge is used in Entree, and other FindMe style systems, to give more accurate similarity assessments 
+of restaurants by considering factors such as price and quality of experience as well as location and cuisine, 
+and by arranging these factors hierarchically. 
+For example, Entree arranges these as: cuisine, price, quality of experience, others, but other systems allow the 
+user to specify this order themselves. 
+This is part of the reason why Entree would outperform this system.
 
@@
 What could I do in the future \cite{5284958}. 
-Implement the idea of particular categories being closer some than others. 
+Implement the idea of particular cuisines being closer some than others. 
 Specify price range and/or consider price when comparing similarities of restaurants. 

### 2021-02-11 RS.py

@@ -15,27 +15,32 @@
 # restaurants
 
 import json
-#from surprise import BaselineOnly
 from surprise import SVD
 from surprise import Dataset
 from surprise import Reader
-#from surprise.model_selection import cross_validate
 import texttable
 import pandas as pd
+from surprise import BaselineOnly
+from surprise.model_selection import cross_validate
+from surprise.model_selection import train_test_split
+from surprise import accuracy
 
 def getRecommendations(uid):
     file_path = 'processed/reviews.csv'
     reader = Reader(line_format='user item rating', sep=',')
     data = Dataset.load_from_file(file_path, reader=reader)
-    #cross_validate(BaselineOnly(), data, verbose=True)
-
-    trainset = data.build_full_trainset()
+    cross_validate(BaselineOnly(), data, verbose=True)
+    #algo = BaselineOnly()
+    #trainset = data.build_full_trainset()
+    #trainset,testset = train_test_split(data, test_size=.25) # for evaluation purposes
     algo = SVD()
-    #start = time.time()
+    cross_validate(algo,data,measures=['RMSE','MAE'], cv=5, verbose=True)
+    exit()
     print('Getting recommendations for {} of {} restaurants in {}...'.format(name,category,city))
     algo.fit(trainset)
-    #end = time.time()
-    #print(end-start)
+    predictions = algo.test(testset)
+    accuracy.mae(predictions)
+    exit()
     suggestions = {}
     for iid in iids:
         if iid in businesses and category in businesses[iid]['categories']:
@@ -43,14 +48,7 @@ def getRecommendations(uid):
             pred = algo.predict(uid,iid)#,verbose=True)#,r_ui=4, verbose=True)
             suggestions[pred[1]] = pred[3]
     suggestions = sorted(suggestions.items(), key=lambda item: item[1],reverse=True)
-    #print(suggestions)
     return suggestions[:8] # present top 8 restaurants
-    #testset = trainset.build_anti_testset()
-    #predictions = algo.test(testset)
-
-    #top_n = get_top_n(predictions, n=8) # get top 8 recommendations
-    #print(top_n.items())
-    #return top_n.items()
 
 def hybrid():
     pass

### 2021-02-12 paper/paper.tex
finished first draft of paper

@@
 The MAE of this system is around 0.99.
+
 Measuring the accuracy of usage predictions is done using precision, calculated by: 
 \begin{equation}
-    \frac{TP}{TP+FP}
+    Precision = \frac{TP}{TP+FP}
 \end{equation}
@@
 The precision of this system is 0.125, since 1 out of the 8 items is a true positive.
-Novelty, diversity, coverage
+
+Coverage of the system is calculated by:
+\begin{equation}
+    Prediction\ coverage = \frac{|I_p|}{|I|}
+\end{equation} 
+where $I_p$ is the set of restaurants which can be recommended and $I$ is the set of all restaurants. 
+In my system the coverage is ~0.32 ($\frac{20,565}{63,944}$), this is the most appropriate metric since 
+the data for previously views items does not exist, ruling out novelty and diversity metrics.
+
+Explainability is measured holistically, by looking at the explanations given throughout the system and how 
+well they describe the purpose of the recommendations, how accurately they match the mechanisms used to 
+generate recommendations, how much they improve the system's transparency, and how they help users to make decisions 
+more efficiently and effectively. 
+Generally the explainability of this system is quite high with regards to all four of these criteria.
 
@@
     \centering
-    \includegraphics[width=0.5\linewidth]{output.png}
+    \includegraphics[width=0.4\linewidth]{output.png}
     \caption{System output}
@@
 \subsection{Ethical issues}
-People's personal data
+One ethical issue resulting from data collection is whether people who use Yelp are aware that some of their 
+personal data is being made publicly available as part of a large dataset. 
+An approach to solve this issue would be an opt-in for whether a user wants to allow their data to be used in this way. 
+Another issue is a business which has received unfair negative reviews, my recommender system would be less likely to 
+suggest this restaurant to a user, causing the owner to unfairly lose business. 
+This can be solved by moderation of reviews and allowing business owners to appeal negative reviews. 
+A third issue arising from this system is that users now don't need to visit the Yelp website to find restaurant 
+recommendations, meaning that Yelp will lose ad-revenue due to reduced site-traffic. 
+Yelp could address this issue by introducing additional useful features to their website which are not available in 
+this system, such that users would want to visit Yelp.com for their recommendations.
 
@@
 \subsection{Limitations}
-What can it not do?
+Currently the system cannot consider the user's preferences in terms of pricing of a restaurant, and it also does 
+not have precise location information, only the metropolitan area the user is in or near to. 
+This means that how central a restaurant is to the city centre currently has no impact on the recommendations.
 
 \subsection{Further developments}
-What could I do in the future \cite{5284958}. 
-Implement the idea of particular cuisines being closer some than others. 
-Specify price range and/or consider price when comparing similarities of restaurants. 
-Include demographic information
+The system could be developed further by implementing the idea of particular cuisines being more similar 
+to some than others. 
+The user could also specify an acceptable price range that they are looking for, and the system could make 
+recommendations based on estimated or known demographic information of the user.
 

### 2021-02-12 filter.py

@@ -310,6 +310,18 @@ def getDemoIDs():
     print(sorted(uids.items(), key=lambda item: item[1]))
     print(len(uids))
 
+def getNumberRestaurants():
+    inFile = open('dataset/business.json','r')
+    count = 0
+    for line in inFile:
+        business = json.loads(line)
+        ID = business['business_id']
+        busCategories = business['categories']
+        if busCategories != None and 'Restaurants' in busCategories:
+            count += 1
+    inFile.close()
+    print(count)
+
 #ids = getIDs()
 #filterRelevant()
 #lastCheckin()
@@ -332,3 +344,4 @@ def getDemoIDs():
 #formatCovidFeatures()
 #closedCovid()
 #getDemoIDs()
+getNumberRestaurants()

### 2021-02-12 RS.py
moved files for video

@@ -25,48 +25,27 @@ from surprise.model_selection import cross_validate
 from surprise.model_selection import train_test_split
 from surprise import accuracy
 
-def getRecommendations(uid):
+def getRecommendations(uid): # 
     file_path = 'processed/reviews.csv'
     reader = Reader(line_format='user item rating', sep=',')
     data = Dataset.load_from_file(file_path, reader=reader)
-    cross_validate(BaselineOnly(), data, verbose=True)
+    #cross_validate(BaselineOnly(), data, verbose=True) # base line comparison
     #algo = BaselineOnly()
     #trainset = data.build_full_trainset()
     #trainset,testset = train_test_split(data, test_size=.25) # for evaluation purposes
-    algo = SVD()
     cross_validate(algo,data,measures=['RMSE','MAE'], cv=5, verbose=True)
-    exit()
+    algo = SVD()
     print('Getting recommendations for {} of {} restaurants in {}...'.format(name,category,city))
     algo.fit(trainset)
-    predictions = algo.test(testset)
-    accuracy.mae(predictions)
-    exit()
     suggestions = {}
     for iid in iids:
-        if iid in businesses and category in businesses[iid]['categories']:
-            #if iid in categories and category in categories[iid]: # include all reviews?
-            pred = algo.predict(uid,iid)#,verbose=True)#,r_ui=4, verbose=True)
-            suggestions[pred[1]] = pred[3]
+        if iid in businesses and category in businesses[iid]['categories']: # knowledge based filtering
+            pred = algo.predict(uid,iid) # predict what user uid will rate item iid
+            suggestions[pred[1]] = pred[3] # collaborative filtering
     suggestions = sorted(suggestions.items(), key=lambda item: item[1],reverse=True)
     return suggestions[:8] # present top 8 restaurants
 
-def hybrid():
-    pass
-    # use namsor to get info about person from name, then use to recommend 
-    # take scores produced by SVD calculations, e.g. when there's nothing to go on
-    # break ties by looking at review count, accepts credit cards, price range, attire, alcohol, 
-    # reservations, takeout, good for groups, ambience, good for kids, drive thru, wifi, parking, 
-    # caters, delivery, noise level, outdoor seating, has tv, good for meal, categories, hours
-    # all from business.json or business_ids.json
-    # TODO
-    # implement hybrid properly, look at lectures to figure out how to do it 
-    # nearest neighbour for e.g. best night or ambience that person likes the most
[…]

### 2021-02-12 filter.py

@@ -344,4 +344,4 @@ def getNumberRestaurants():
 #formatCovidFeatures()
 #closedCovid()
 #getDemoIDs()
-getNumberRestaurants()
+#getNumberRestaurants()

### 2021-02-12 RS.py
fixed uncommented test code

@@ -30,10 +30,10 @@ def getRecommendations(uid): #
     reader = Reader(line_format='user item rating', sep=',')
     data = Dataset.load_from_file(file_path, reader=reader)
     #cross_validate(BaselineOnly(), data, verbose=True) # base line comparison
+    #cross_validate(SVD(),data, verbose=True)
     #algo = BaselineOnly()
     #trainset = data.build_full_trainset()
     #trainset,testset = train_test_split(data, test_size=.25) # for evaluation purposes
-    cross_validate(algo,data,measures=['RMSE','MAE'], cv=5, verbose=True)
     algo = SVD()
     print('Getting recommendations for {} of {} restaurants in {}...'.format(name,category,city))
     algo.fit(trainset)

### 2021-02-12 RS.py
renamed rs.py and uncommented necessary code

@@ -32,10 +32,10 @@ def getRecommendations(uid): #
     #cross_validate(BaselineOnly(), data, verbose=True) # base line comparison
     #cross_validate(SVD(),data, verbose=True)
     #algo = BaselineOnly()
-    #trainset = data.build_full_trainset()
     #trainset,testset = train_test_split(data, test_size=.25) # for evaluation purposes
     algo = SVD()
     print('Getting recommendations for {} of {} restaurants in {}...'.format(name,category,city))
+    trainset = data.build_full_trainset()
     algo.fit(trainset)
     suggestions = {}
     for iid in iids:

### 2021-02-12 paper/paper.tex
spelling corrections to paper

@@
 that the review is written for. 
-The file $user.json$ holds every user's first name along with other data adn metadata which I don't use \cite{YelpDataset}. 
+The file $user.json$ holds every user's first name along with other data and metadata which I don't use \cite{YelpDataset}. 
 The data files $checkin.json$ and $tip.json$ are not used in my system. 
@@
 and the collaborative-filtering system as the secondary the hybrid system can find hidden similarities between 
-restaurants in different cities very accurately. 
+restaurants in different cities accurately. 
 This method of hybrid recommender system (HRS) has been implemented as a restaurant recommender in the Entree 
@@
 This approach was chosen because it scales well with large datasets, which is important since my dataset has 
-~120,000 entries. 
+roughly 120,000 entries. 
 The CF technique here also introduces an element of variability in the recommendations, such that the user 
@@
 In my system the coverage is ~0.32 ($\frac{20,565}{63,944}$), this is the most appropriate metric since 
-the data for previously views items does not exist, ruling out novelty and diversity metrics.
+the data for previously viewed items does not exist, ruling out novelty and diversity metrics.
 
@@
 \subsection{Input interface}
-The system has a command line based interface, so all user input of the input is given via the command line. 
-The system recognises the active user by their unique user ID, some example IDs are provided in the 'README.txt' 
-file for testing and demonstration purposes. 
+The system has a command line based interface, so all user input is given via the command line. 
+The system recognises the active user by their unique user ID. 
+Some example IDs are provided in the 'README.txt' file for testing and demonstration purposes. 
 Only explicit user data is gathered in order to make the system more explainable and transparent, 
@@
 and by arranging these factors hierarchically. 
-For example, Entree arranges these as: cuisine, price, quality of experience, others, but other systems allow the 
-user to specify this order themselves. 
+For example, Entree arranges these as: cuisine, price, and quality of experience, in descending order of importance 
+and other systems allow the user to specify this order themselves. 
 This is part of the reason why Entree would outperform this system.

### 2021-02-12 paper/paper.tex
finished

@@
 \IEEEauthorblockA{\textit{dept. Computer Science} \\
-\textit{Durham University}\\
+\textit{Durham University} \\
 Durham, United Kingdom}
@@
 Hybrid recommender systems which combine knowledge-based (KB) and collaborative-filtering (CF) techniques 
-are designed to combine the ability of KB techniques to give generic recommendations to new users with 
-small profiles and the strength of CF to find peer users with unexpected shared preferences. 
-This combination of techniques is particularly well suited to restaurants since a user may often be in the 
-mood for a specific type of food, such as Italian or Chinese, so they only want to see suggestions of that 
-type or restaurant, which is where the knowledge-based recommender is helpful. 
-A collaborative-filtering algorithm will then be able to order all the restaurants which match this 
-explicit criteria by how the user should rate them, based on the opinions of other similar users, 
-helping them to discover new restaurants to try. 
+are designed to combine the ability of KB techniques to give appropriate recommendations to new users with 
+the ability of CF to find peer users with unexpected shared preferences. 
+This is appropriate for restaurants since a user may be in the mood for a specific cuisine and only want to see 
+suggestions relevant to that, which is where the KB system comes in. 
+The CF algorithm will then be able to sort these restaurants provided by the KB system by their predicted 
+rating by the user, based on the opinions of other similar users, helping them to discover new restaurants to try. 
 These techniques are used in the Entree restaurant recommender system \cite{burke1999integrating,burke2000knowledge,burke2002hybrid,burke2007hybrid}. 
@@
 This is a subset of all the businesses, reviews, and user data on Yelp, formatted as JSON files. 
-The dataset includes approximately 8,000,000 reviews of 200,000 businesses by 2,000,000 users, as well as 
-additional Covid-19 related features of each business such as whether they now offer takeout or delivery. 
-There are 3 sections of the dataset, with 5 JSON files in the main dataset, 1 in the photos section, and 
-1 in the covid section. 
 The $business.json$ file contains business data including location data, attributes, and cuisine. 
-$review.json$ contains full review text data including the $user\_id$ that wrote the review and the $business\_id$ 
-that the review is written for. 
-The file $user.json$ holds every user's first name along with other data and metadata which I don't use \cite{YelpDataset}. 
-The data files $checkin.json$ and $tip.json$ are not used in my system. 
-The dataset also includes approximately 200,000 photos but these are not relevant to my application, since 
-the interface is command line only, so they were discarded. 
+$review.json$ contains full review text data including the $user\_id$ of the review and the $business\_id$ 
+that the review is for. 
+The file $user.json$ holds every user's first name along with other data and metadata which I don't 
+require \cite{YelpDataset}. 
+The data files $checkin.json$, $tip.json$, and $photos.json$ are not used in my system. 
 The covid related data from $covid\_features.json$ contains additional information about each business with regards to 
@@
 special message for customers. 
-Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada 
-(Montreal, Calgary, Toronto, Pittsburgh, Charlotte, Urbana-Champaign, Phoenix, Las Vegas, Madison, and Cleveland) \cite{YelpDataset}.
+Every business in the dataset is in 1 of 10 metropolitan areas in the USA and Canada \cite{YelpDataset}.
 
 \subsection{Data preparation and feature selection}
-The entire Yelp dataset contains a lot of data which is not relevant to my domain of restaurants, so much data 
-preparation is required to save computation time and in order to provide more accurate predictions. 
 The dataset is provided in JSON format, but in such a way that every entry is its own JSON on its own line, 
 which does not lend well to fast lookups by the main program. 
-Therefore it was necessary to convert every file I used in the system into a typical JSON file, such that it could 
-be loaded as a python dictionary with the lookup key being either the uniquely identifying $business\_id$ or 
-$user\_id$ depending on the case. 
-This allows for extremely fast data lookups in the system. 
-It was of course also necessary to filter the list of all Yelp businesses down to just the restaurants. 
-I further narrowed the list of restaurants by selecting only those which fit into one or more of the top 12 
-cuisines in terms of number of restaurants (American, Mexican, Italian, Chinese, Seafood, Japanese, Canadian, 
-Mediterranean, Indian, Thai, Middle Eastern, Vietnamese, in that order). 
-Since the user selects one of these 12 options for cuisines they want to find, they will 
-never be recommended a restaurant which doesn't fit at least one of these categories. 
-I chose to use the top 12 cuisines because it's enough to provide a wide variety of options to the user without 
-being so many that the user is overwhelmed by the amount. 
-I also ensured that every restaurant was still open and had at least 1 check-in in 2019 (the last year that 
-data was available). 
+So I converted every file I used in the system into a typical JSON file, allowing for extremely fast 
+data lookups in the system. 
+I filtered the list of all Yelp businesses to restaurants and then further by selecting only those which were still 
+open and fit into one or more of the top 12 cuisines in terms of number of restaurants because it's enough to provide 
+a wide variety of options to the user without being so many that the user is overwhelmed. 
 This left 20,565 restaurants which could be recommended to the user, which needed to be split into their respective 
-metropolitan areas by using the latitude and longitude values of the businesses and of the centres of each city. 
-All other datasets were filtered to only data pertaining to these restaurants, with duplicates in the 
-covid dataset removed. 
-The reviews were further filtered to only those which had 5 or more 'useful' votes by other users. 
-This was to ensure I was only using high quality reviews which others found helpful, as well as to allow the 
-recommender system to compute results in a reasonable amount of time (~10 seconds on my machine). 
-This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) with unnecessary data 
-removed and structured as follows: ($User\_ID$, $Business\_ID$, $Rating$).
+metropolitan areas. 
+The reviews were further filtered to only those which had 5 or more 'useful' votes by other users, 
+to ensure I was only using high quality reviews which others found helpful. 
+This left around 120,000 reviews, which were then converted to a CSV file ($reviews.csv$) and structured as 
+follows: ($User\_ID$, $Business\_ID$, $Rating$).
 
@@
 matrix factorisation. 
-This approach was chosen because it scales well with large datasets, which is important since my dataset has 
-roughly 120,000 entries. 
+This approach was chosen because it scales well with large datasets. 
 The CF technique here also introduces an element of variability in the recommendations, such that the user 
@@
 due to some rare skewed ratings.
-The MAE of this system is around 0.99.
 
@@
 \subsection{Input interface}
-The system has a command line based interface, so all user input is given via the command line. 
-The system recognises the active user by their unique user ID. 
-Some example IDs are provided in the 'README.txt' file for testing and demonstration purposes. 
-Only explicit user data is gathered in order to make the system more explainable and transparent, 
+The system has a command line interface, and recognises the active user by their unique user ID. 
+Some example IDs are provided for testing and demonstration purposes. 
+Only explicit user data is gathered, making the system more explainable and transparent, 
 and users are made aware of which data is collected, how it is collected, and for which purposes. 
@@
 When the user is presented with a list of recommendations, they can then choose one that they want 
-to see more information about and from here they can leave a rating, return to the list, or return to 
-the main menu. 
-User profiles can be updated by giving a rating of a restaurant through the system. 
-Any rating given will affect future recommendations given by the system to that user 
-and to other similar users.
+to see more information about and leave a rating, which will update their future recommendations.
 
@@
 \end{equation}
-Where the parameters are learnt by minimising the following regularised squared error: 
+Where the parameters are learnt by minimising the following regularised squared error by stochastic gradient descent: 
 \begin{equation}
@@
 \end{equation}
-This minimisation is performed by stochastic gradient descent:
-\begin{align}
-    \begin{split}
-    & b_u \gets b_u + \gamma(e_{ui} - \lambda b_u) \\
-    & b_i \gets b_i + \gamma(e_{ui} - \lambda b_i) \\ 
-    & p_u \gets p_u + \gamma(e_{ui} \cdot q_i - \lambda p_u) \\ 
-    & q_i \gets q_i + \gamma(e_{ui} \cdot p_u - \lambda q_i) \\
-    \end{split}
-\end{align}
-where $e_{ui} = r_{ui} - \widehat{r}_{ui}$. 
-These steps are performed over all the ratings of the trainset and repeated 20 times \cite{SVD}. 
 The knowledge-based aspect of the algorithm produces recommendations based on the explicit location and cuisine 
@@
 \subsection{Output interface}
-Recommendations and prediction scores are presented to the user via the command line interface and using an ASCII table. 
-I used a python module called texttable for this, since it makes the output very clear and visually appealing. 
-I chose to output the top 8 recommendations since I think this is enough variety and choice for the user to find 
-a restaurant to try, without there being too many options which might overwhelm the user. 
-If I presented more than 8 recommendations at a time, the table wouldn't fit in a single terminal output window and 
-would require the user to scroll around to read different parts of the output. 
+Recommendations and prediction scores are presented to the user via the command line interface using an ASCII table. 
+The top 8 recommendations are output since this is enough variety for the user to find something new to try, 
+without being too overwhelming. 
 Since the user will be interacting with the system on a laptop or desktop computer through a command line interface, 
@@
 well as the prediction values themselves so the user can understand why they are being shown these suggestions. 
-This, along with an option in the main menu to view information about the data the system uses and how it is used, 
+This, along with an option in the main menu to view information about the data the system collects and uses, 
 further adds to the explainability and transparency. 

### 2021-02-12 RS.py

@@ -1,48 +1,35 @@
 
-# you must compare the other systems using evaluation metrics but actual performance of recommender system is not important
-
-# video is about how well you can present your product. Voice over video
-# explain what it is, how it works, display the best of its features, demo main functionality
-# like a little sales pitch
-
-# when reviewing papers, look at:
-#   RS technique/type
-#   model
-#   evaluation methods used
-#   limitations
-# in order to get an idea of common or good or useful combinations of systems
-
-# restaurants
+# Restaurants
+# knowledge-based and collaborative-filtering in cascade style
 
 import json
+import texttable
 from surprise import SVD
 from surprise import Dataset
 from surprise import Reader
-import texttable
-import pandas as pd
-from surprise import BaselineOnly
-from surprise.model_selection import cross_validate
-from surprise.model_selection import train_test_split
-from surprise import accuracy
+#from surprise import BaselineOnly # testing import
+#from surprise.model_selection import cross_validate # testing import
+#from surprise.model_selection import train_test_split # testing import
+#from surprise import accuracy # testing import
 
-def getRecommendations(uid): # 
-    file_path = 'processed/reviews.csv'
-    reader = Reader(line_format='user item rating', sep=',')
+def getRecommendations(uid): # hybrid recommender
+    file_path = 'processed/reviews.csv' # path to ratings file
+    reader = Reader(line_format='user item rating', sep=',') # specify line order of CSV file
     data = Dataset.load_from_file(file_path, reader=reader)
     #cross_validate(BaselineOnly(), data, verbose=True) # base line comparison
     #cross_validate(SVD(),data, verbose=True)
     #algo = BaselineOnly()
     #trainset,testset = train_test_split(data, test_size=.25) # for evaluation purposes
-    algo = SVD()
+    algo = SVD() # Single Value Decomposition
     print('Getting recommendations for {} of {} restaurants in {}...'.format(name,category,city))
-    trainset = data.build_full_trainset()
-    algo.fit(trainset)
+    trainset = data.build_full_trainset() # training data
+    algo.fit(trainset) # create model
     suggestions = {}
-    for iid in iids:
-        if iid in businesses and category in businesses[iid]['categories']: # knowledge based filtering
+    for iid in iids: # iid is business ID
+        if iid in businesses and category in businesses[iid]['categories']: # knowledge based filtering, only predict items recommended by KB system
             pred = algo.predict(uid,iid) # predict what user uid will rate item iid
[…]

### 2021-02-12 filter.py

@@ -2,7 +2,6 @@
 import json
 import pandas as pd
 
-
 def getIDs():
     inFile = open('processed/2019business.json','r')
     outFile = open('processed/business_ids.json','w')
